### 2021-03-01 main.tex
Cleaned up formalism

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2021}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\input{macros}
\input{examples}

\begin{document}

%% Title information
\title{Stack Safety as a Security Property}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
%\author{First1 Last1}
%\authornote{with author1 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%\affiliation{
%  \position{Position1}
%  \department{Department1}              %% \department is recommended
%  \institution{Institution1}            %% \institution is required
%  \streetaddress{Street1 Address1}
%  \city{City1}
%  \state{State1}
%  \postcode{Post-Code1}
%  \country{Country1}                    %% \country is recommended
%}
%\email{first1.last1@inst1.edu}          %% \email is recommended
%
%%% Author with two affiliations and emails.
%\author{First2 Last2}
%\authornote{with author2 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%\affiliation{
%  \position{Position2a}
%  \department{Department2a}             %% \department is recommended
%  \institution{Institution2a}           %% \institution is required
%  \streetaddress{Street2a Address2a}
%  \city{City2a}
%  \state{State2a}
%  \postcode{Post-Code2a}
%  \country{Country2a}                   %% \country is recommended
%}
%\email{first2.last2@inst2a.com}         %% \email is recommended
%\affiliation{
%  \position{Position2b}
%  \department{Department2b}             %% \department is recommended
%  \institution{Institution2b}           %% \institution is required
%  \streetaddress{Street3b Address2b}
%  \city{City2b}
%  \state{State2b}
%  \postcode{Post-Code2b}
%  \country{Country2b}                   %% \country is recommended
%}
%\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
What is stack safety? Numerous mechanisms have been proposed to guard
against attacks that corrupt the call stacks of running programs,
\ifaftersubmission
and many
others rely on a protected stack as an assumption to provide higher level
guarantees,\bcp{this is a good point, but we don't pick it up anywhere in the
  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}
\bcp{Let's delete for now; later, we can try to incorporate it in the body
  of the paper.}
\fi
but a clear and precise definition of stack safety is missing
from the literature.
%
We propose a formal definition phrased in the technical
terminology of language-based security---as a combination of a
confidentiality property (callees should not be able to read their caller's
stack frames) and an integrity property (callees should not corrupt their
caller's stack frames or control state), together with a straightforward
well-bracketedness condition on the program's control flow.
  %
We discuss how these properties are enforced or approximated by
existing stack protection mechanisms and describe useful
stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
\leo{Not sure. We almost never cite stuff in the abstract, do we?}\fi
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~General programming languages}
%\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{Stack Safety, Micropolicies}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

\newcommand*{\MemoryLabel}[3]{\raisebox{#2}{\makebox(0,0){\hspace{#1}#3}}}

%\begin{figure}
%  \begin{minipage}{\textwidth}
%    \begin{center}
%\MemoryLabel{10em}{1.5em}{40}
%\MemoryLabel{16em}{1.5em}{60}
%\MemoryLabel{22em}{1.5em}{80}
%$\cdots$
%\memory{4}{gray}%
%\memory{5}{green}[{\makebox[0pt]{Caller}}]%
%\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
%\memory{8}{red}[{\makebox[0pt]{Callee}}]%
%\memory{4}{\unusedc}%
%~$\cdots$\\
%  \end{center}
%  \end{minipage}\\
%~\\
%~\\
%  \begin{minipage}{0.45\textwidth}
%    \begin{center}
%\begin{verbatim}
%jal ...
%lw r1 r0 40
%\end{verbatim}
%    \end{center}
%  \end{minipage}~
%  \begin{minipage}{0.45\textwidth}
%    \begin{center}
%\begin{verbatim}
%jal ...
%sw r0 r0 40
%\end{verbatim}
%    \end{center}
%  \end{minipage}
%  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
%      the boxes; the program fragments need comments explaining what each
%      line does.}}
%  \label{fig:stackunsafety}
%\end{figure}
%
%\leo{I hate the writing here, but I wanted to make a start.}

%\bcp{We don't actually say very clearly anywhere that we are interested in
%  low-level enforcement mechanisms for machine-code programs (because this
%  is where stack attacks occur, and because at higher levels of abstraction
%  the stack is a built-in concept, not something that can be attacked even
%  in principle).  This could cause confusion.\apt{not so worried}}
%\leo{I agree. I've clarified the writing in the top of the running example
%  a bit to make sure though}

The call stack is a perennial target for attacks
based on memory-safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi

But what {\em is} this stack safety of which people speak? Precise formal
definitions are mostly lacking: rather, the notion
is commonly defined in a negative way, by enumerating the bad things that a
given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a
positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as
the conjunction of two properties: {\em local state encapsulation} and {\em
  well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft an operational characterization
of stack safety that is independent of capabilities or any other specific
enforcement technique.  In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.''
%
Our key contribution is an abstract presentation of local state encapsulation
using the technical framework of language-based security, together with a simple
characterization of well-bracketed control flow.

\begin{figure}[b]
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 42;
  f();
  print(x);
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 1;
  f();
  if(x)
    g();
  else
    h();
}
\end{verbatim}
    \end{minipage}
  \caption{Simple Examples}
  \label{fig:examples1}
\end{figure}

\begin{figure}[b]
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main(int arg) {
  if(arg)
    g();
  else
    h();
  f();
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.4\textwidth}
\begin{verbatim}
int main() {
  int x = 1;
  f();
  x = -x;
  f();
  print(x);
}
\end{verbatim}
  \end{minipage}

  \caption{More Examples}
  \label{fig:examples2}
\end{figure}

\paragraph*{The Use of Stack Safety}

When does the programmer rely on stack safety? The most natural example is the first
in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
overwrite it's caller's stack data. To this end, our properties support a
{\em call rule} that can be used to reason that on return from {\tt f}, {\tt x} still
contains the same value. This is not only useful for the programmer: consider the second
example, where the call to {\tt h} should be unreachable and therefore subject to dead code
elimination, if only the compiler can soundly reason that {\tt x} does not change.
The property that enables the call rule is, in security terms, {\em integrity}: callers'
data is protected from modification by callees.

Alongside integrity, we have {\em confidentiality}, in which callers' data is protected
from leakage by their callees. Consider again the first example in figure \ref{fig:examples1},
in which {\tt f}'s behavior should not depend on {\tt x} being 42. In fact suppose that the
programmer had initialized {\tt x} after calling {\tt f} rather than before; such a change
should not alter the behavior of {\tt main}, but if {\tt f} can directly read {\tt x}, it
may. But there are more subtle examples, such as the first in figure \ref{fig:examples2},
where {\tt main} may call {\tt g} or {\tt h} are called prior to calling {\tt f}, depending on
its argument. Here, if {\tt g} or {\tt h} leaves behind identifying information in unused stack
space, {\tt f} could still deduce information about the argument and behave differently, violating
the abstraction of its call. So our confidentiality properties also support a {\em insensitivity rule},
that each function's behavior is independent of the state of its caller, other than arguments.
This also makes it easier, though not trivial, to prove a function's idempotence: the channels through
which a callee can remember its own history are limited to those outside the stack.

Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
{\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
callers from their callees. Consider the last example in figure \ref{fig:examples2},
a variant of the famous ``awkward example,'' in which {\tt f} is called twice and may not
be idempotent. It should be the case that {\tt main} prints -1, but if
{\tt f}'s first activation stores its return address somewhere accessible, the second
activation can return to before {\tt x} is negated, causing {\tt x} to be set to
1 again before a third activation that would return normally. Then {\tt main} would print 1.
Once again this is rather unexpected behavior.

Preventing this form of reentrancy attack is the intuition behind
\citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
the call graph itself. A callee should pass control back to the caller only by
transferring control to the instruction following the call; it should not be allowed
to jump back into the caller's code at an arbitrary point, either directly (via a jump
instruction) or indirectly (by corrupting stack data and returning).

To phrase LSE and WBCF as security properties requires an explicit attacker model:
a clear understanding of who is being protected and from what. Most existing
stack protection mechanisms envisage scenarios where a callee is
influenced by external inputs to corrupt the stack in some way. To
avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we
state the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns.

A key challenge in adapting the concepts of language-based security to this
setting is that integrity and confidentiality are ``nested'' properties: not
only does a given caller need to be protected from its callee, but the callee
also needs to be protected from {\em its} callees, and so on. That is, the notion
of ``attacker'' changes at many points throughout the execution of the program.
We address this issue by introducing a dynamically model of {\em domains} for
the different components of the machine state, in which some stack locations are
sealed by the caller that needs to preserve them while the rest are unsealed and
available for use.

First we introduce this model for a single simple stack with no sharing between callers and
callee. Then we expand it to allow components to be passed, granting a callee one time
access to the caller's data, or shared, allowing components to be outright
exempted from the stack safety properties. Shared components could model objects whose
addresses are passed and whose security must be modeled by a separate property, such as
memory safety enforced by capabilities.

After formally defining stack safety, we show that the definition is
enforceable using an existing tag-based enforcement
mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
\citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy. 
We give {\em eager} properties that this policy enforces, in which each
callee, at the point of its return, can be seen to respect the integrity
and confidentiality of its caller. Additionally we have an
{\em ultra-eager} integrity property which guarantees that integrity
is respected at each step. The advantage of this property is that it ``fails
faster'' and is therefore more suitable for efficient testing. We prove (in
Coq) that the ultra-eager property implies our original definition of stack
safety.

The downside of the Depth Isolation policy is that it is rather
expensive to implement. In the same work,
\citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
that is more efficient by virtue of being lazy in its enforcement.
Instead of enforcing the integrity and confidentiality of the stack at the
exact moment when a callee returns, this policy signals violations when the
caller actually accesses data that the callee has written into its stack
frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
  property in the same way...}\fi.
The lazy policy does not enforce our original integrity and confidentiality
properties, and to our knowledge there is no formal statement of the properties that
it ought to enforce. We propose a {\em lazy} integrity property, which capture the
desired behavior of a policy that detects violations between when they occur and when
they make a visible impact. Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s
existing ``Lazy Tagging and Clearing'' policy does not quite enforce these properties,
but formalizing the ideal property allows us to see how it falls short.
We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
at some performance cost.

We extend our properties to allow sharing, both in the form of arguments passed
on the stack and of objects shared widely beyond their owning function. In the latter
case, restricting access the shared objects is beyond the scope of stack safety,
but our model delimits them from other objects in the same frame.

%\bcp{I worry that we are being a little
%  dishonest here: the stated advantage of the lazy policy is that it is more
%efficient, but our variant is {\em not} efficient!}

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
  A novel {\em stack-safety property} abstractly capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}).
  % These properties rule out common examples of stack memory abuse.
  This property can be enforced by
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
  based on hardware tagging~(\cref{sec:conservative}).
\item
  A {\em lockstep} variant of the stack-safety property that is also
  enforced by the Depth Isolation policy
  but is more efficient to test, plus a proof (in Coq) that this variant is
  stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced if the
  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
  with unique tags on each activation, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.

We begin with a description of our threat model
(\cref{sec:threat}), a running example that we refer to throughout the paper
(\cref{sec:running-example}),  and some technical preliminaries
(\cref{sec:prelim}).

\section{Threat Model and Assumptions}
\label{sec:threat}

Stack safety is meant to protect a caller from a confused or malicious callee's
attempts to access private data or hijack its control flow. Our
attacker model reflects this intention by assuming that the attacker
can run arbitrary machine code to try to access or overwrite its caller's stack data
or disrupt its control state.  Hardware and timing attacks are out of
scope.
%\leo{Anything else to add here?}
%\leo{Maybe a good place to discuss calling conventions and that they
%  are essentially enforced by the policy or sth?}
%\rb{Intuition that the callee can do anything in principle, but the machine will
%halt if it does not follow the calling convention, etc.?}

\newcommand*{\rsp}{\textsc{sp}}

We assume we are given a machine language program, together with some auxiliary information
about its structure: the code footprint of each function and the locations of instructions
representing calls, returns, and function entry points.
Although this information would typically come from a compiler together with
the machine code, we do not
assume anything about their provenance.  We assume that the machine has a
program counter $\PCname$
and stack pointer $\rsp$; other details of the machine's ISA are unimportant.

To make our stack safety definitions precise, we make some straightforward
assumptions about
memory layout; our definitions could easily be changed to be parametric over
these assumptions, at the cost of some complexity in their statements.  Memory
consists of an instruction region and a stack that grows upwards. (For simplicity,
we do not consider the heap. Integrating a memory-safety property for the
heap in the style of
\citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
in~\Cref{sec:relwork}.)  The stack has a conventional structure, with
contiguously allocated per-function frames.
The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
at the time of the call.  The stack can be used to
pass arguments or return results in the usual way, by placing them
at the top of the caller's frame. Callers and callees can also communicate
via registers if they wish; indeed, our
definitions say nothing at all about registers (except $\PCname$ and
$\rsp$). \ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi

Our stack-safety concepts depend critically on the notions of ``call'' and ``return,''
which are not fully explicit in machine code (the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes).  We define a call to be
any instruction declared as such in the compiler-provided call structure.
If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
  return} has occurred when the $\PCname$ points to the next instruction after $a$ and $\rsp = s$. (Our definition
of well-bracketed control flow relates this definition of return to the program's
declared set of return points.)
Our property definitions make no other detailed assumptions about call or return sequences, although
some of the enforcement mechanisms we consider do.

Enforcement mechanisms are modeled as {\em policies} (borrowing the
terminology from tag-based mechanisms) that add some auxiliary policy
state to the ordinary machine state
%\bcp{This may be surprising to people who are not thinking about micropolicies.\apt{added para below}}
and a partial policy step function that refines the machine step
function by allowing or disallowing a step to be taken based on its
auxiliary state.  Enforcement works by halting the program before it
can perform an action that would violate the security property; our
properties are therefore naturally \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

We model ordinary program state and policy-enforcement state separately to make it
easier to use well-established concepts from language-based security.
In particular, confidentiality has an elegant extensional characterization in terms of
\emph{noninterference}: a program preserves the confidentiality of certain data if
\emph{varying} that data would not change the program's observable
behavior~\citep{6234468}.
In applying this idea to a system that incorporates policy enforcement, only
ordinary program data should be varied, while any associated policy
state should be left alone; our model makes this distinction obvious.
%That is, we compare the actual execution of the program with a \emph{hypothetical}
%execution where data has been varied.
We will also introduce a notion of {\em observational integrity} that relies on the idea of
\emph{rolling back} program state; again, our model makes it easy to describe
rollback of ordinary program data while leaving policy state alone.

%% \leo{Policies of interest: null, micro, cherri, software-only}
%% \leo{TODO: Write down carefully}

This enforcement model directly describes
micropolicies~\citep{pump_oakland2015}, and we believe that it is
flexible enough to capture code-altering approaches (such as code rewriting
or capability-based techniques).

\newcommand{\TAGS}{\mathcal{T}}
\newcommand{\tagname}{t}
\newcommand{\uP}{{\mu P}}

\paragraph*{Micropolicies}
%
Micropolicies are a flexible tag-based, hardware-accelerated
reference monitoring mechanism; they been applied to stack safety
enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
Here the policy state consists of metadata tags (e.g., identifying stack frames)
attached to each value in memory or registers, and the policy step
function checks that each machine operation obeys a set of rules on tags
(e.g., that the current $\PCname$ tag matches the tag on the stack
location being accessed), and halts the machine if not.  Distinguishing
ordinary machine state and policy state is very natural in this setting.
%% Here the machine state is extended with tags paired with all components of the
%% system, and a monitor checks for each operation that the tags on the operands
%% obey a set of rules. One could treat tags directly as part of the machine
%% state, and build the rules into the step function. But local state
%% encapsulation properties as we frame them are not intuitive in their
%% interaction with tags or other additional enforcement state, so it is helpful
%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
%% a software fail state.
%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
%%   policy? }

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines such as CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including
the stack protection policies of
\citet{Skorstengaard+19b} and (with an extension to support linear capabilities) \citet{Skorstengaard+19}.
%\bcp{I thought these policies used
%  linear capabilities?}\leo{Only the second one AFAIK. What would you
%  change to characterize it more accurately?}\rb{StkPointers uses linear
%  capabilities, its predecessor users local capabilities (and unlike
%  StkPointers could be implemented in CHERI)}.
Capabilities package up pointers
(ordinary data) with base and bounds information (policy data),
and they require rewriting source code (e.g., to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate; we conjecture it can be done with some effort.

%% Mapping the underlying protection
%% mechanisms to policy states and step function would be a delicate operation, as
%% those are tightly coupled to the hardware design. However, separating the
%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
%% would encompass details like the abstract call stack, the capability sealing
%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
%% the machine would be relatively straightforward.  \bcp{Not sure that any of
%%   this is comprehensible... :-(}

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the
compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
to insert stack canaries~\citep{Cowan+98}. Such approaches can, in principle,
be evaluated in terms of our formal notion of stack safety with a trivial null
policy. But any state maintained by the enforcement mechanism must be in normal
memory and subject to the same confidentiality and integrity concerns (to be
varied, rolled back, etc.), a heavy constraint. Alternately it can be stripped
out as policy state provided that it can be proven to protect itself.

\ifaftersubmission\apt{This still seems wrong to me. The policy part needs somehow to describe everything that isn't varied or rolled back.}\fi
%These approaches are represented as policies that
%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
%being produced by the compiler.) The set of policy states is again the unit type
%and the step function constant. Property enforcement comes from the fact that
%only ill-formed states can induce a trace that violates the property.

%% \Paragraph*{Null policy}
%% %
%% The simplest possible policy is the neutral policy that carries no information
%% and does nothing. Its set of policy states is the unit type, and its policy step
%% function is the constant unit function, thereby allowing all calls to the machine
%% step function. Its initialization function is defined on all states. This
%% corresponds to the lifting of our machine model to the policy setting. Machines
%% that fulfill a property by construction will also fulfill that property with the
%% null policy.

%
%\leo{TODO: Add citations of mitigation techniques}
%\rb{And for those techniques, explain how they reflect their motivating
%examples. What does each technique do for those examples? How are they
%connected to this work?}
%
%\paragraph*{Bad Write}
%
%\[
%44 : [\rsp - 5] \leftarrow r_4
%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
%\]
%
%Instruction 44 currently is
%$ [\rsp - 2] \leftarrow r_4$,
%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
%with the
%effect of storing the contents of $r_4$ as the result of {\tt g}. If
%instead it wrote to location $\rsp - 5$ then it would be overwriting
%private data (the return address) belonging to {\tt f}.
%
%\paragraph*{Bad Read}
%
%\[
%30 : r_5  \leftarrow [\rsp - 6]
%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
%\]
%
%Instruction 30 currently is
%$ r_5 \leftarrow [\rsp - 4]$,
%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
%with the
%effect of loading the argument of {\tt f} to $r_5$. If instead it
%loaded the contents of location $\rsp - 6$, it would obtain access
%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
%outputs the contents of $r_5$, that value would immediately be made public.
%
%\paragraph*{Bad Control Flow}
%
%\[
%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
%%5 : \jalr ~ r_4 ~ \rra ~ 0
%\]
%If a {\jalr} instruction was added to {\tt main} with its target being
%the middle of the code block of {\tt g}.\apt{describe result of this?}

\section{Examples and Intuition}
\input{example1}

\section{Machines, Traces, and Observations}
\label{sec:prelim}

We next describe our machine model. To make our definition of stack safety
as generic as possible, this section proposes an abstract interface to a
machine model plus some kind of (software- or hardware-enforced) policy monitor.
In~\cref{sec:enforcement}, we will instantiate this model with a concrete dynamic
policy enforcement mechanism based on \citet{DBLP:conf/sp/RoesslerD18}.

\subsection{Values, States, and Observations}

The basic building blocks of the machine are {\em values} and {\em
  addresses}. Both are drawn from some set of {\em words}
\(\WORDS\), ranged over by \(\word\).
%
Our machine states are composed of {\em components} \(\component\), addressed either by
words (memory addresses) or by register names (\(\reg\)) drawn from some set
\(\REGS\).
%
    \[\component \in \COMPONENTS = \WORDS + \REGS \]
%
The register names are assumed to include two special purpose registers: the
program counter {\PCname} and the stack pointer \(\SP\).

A {\em machine state} is a map from components to values:
%
\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]

The step function of the machine takes a machine state and produces a
pair of a machine state and an {\em observation}, which may be either
silent (written \(\tau\)) or a word of data\ifaftersubmission\bcp{This is a rather concrete
  choice!  Why not leave the set of observations abstract?}\fi that the machine
communicates to the outside world:
%
\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
\[\mach \stepstoobs[\obs] \mach' \in \MACHS \rightarrow \MACHS \times \OBSS \]
%
This step function is total, so the machine's execution traces
(defined in \cref{sec:traces}) are infinite.

We choose to model observations explicitly, rather than simply treating
a subset of registers or memory as being observable.  This choice gives
us the flexibility to treat programs as observationally equivalent even when
their internal states may differ.
%The choice of modelling observations is an interesting one. A first approach
%would be to consider any change to the registers or memory as observable.
%However, as some parts of the stack will be considered ``private''
%this choice won't work. A relaxation of this to consider all changes
%to ``non-private'' data as observable could work in principle, but
%does not allow for an interesting weaker-but-more-efficient enforcement
%mechanism. We will come back to this point in \Cref{sec:lazy}.

In the running example of \Cref{sec:running-example},
$\WORDS$ is the
set of 64-bit integers and $\REGS$ is the set $\{r_0, \cdots, r_{31}\}
\cup \{\PCname,\SP\}$. Observations are writes to the
special $\rout$ location\ifaftersubmission\bcp{Which is 100! (do we need a name for it? seems
  not...)}\fi, which can be thought of as a memory-mapped
  output port: if during an execution step a word $w$ is
written to $\rout$ then $\obs = w$, otherwise $\obs = \tau$.

\subsection{Policies}

We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
A policy is a general model of an enforcement mechanism, consisting of a set of policy
states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
it is undefined on input configurations that correspond to a policy fault. The policy
can see the machine state, but cannot modify it. We assume that the
compiler provides an initial policy state \(\pol_0\) to accompany the initial machine state.
%
A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
\Cref{sec:enforcement}.

We write $\MPS$ for the set of all pairs of machine states and policy
states.
%
\[\mpstatename \in \MPS = \MACHS \times \POLS\]
%
We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.

\judgmenttwo{\(\mach \stepstoobs[\obs] \mach'\)}{\((\mach, \pol)
               \polstep \pol'\)}
            {\(\mpstate{\mach}{\pol} \stepstopol[\obs]
               \mpstate{\mach'}{\pol'}\)}

This combined step function is also partial, and traces of the machine-with-policy
will be finite iff a policy fault occurs.

\subsection{Contexts}

Separate from the machine state, our formal properties will keep track of additional context
information representing the history of the run. The following definitions are parameterized
over an arbitrary set of context states, \(\CONTEXTS\), an initial context state \(\context_0
\in \CONTEXTS\), and a context step function \(\mach, \context \constep \context' \in
\MACHS \times \CONTEXTS \rightarrow \CONTEXTS\).

We lift contexts into an MP trace in the same way as we lifted policy states to get
a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).

\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol{\obs} \mpstate{\mach'}{\pol'}\)}
            {\(\mach, \context \constep \context'\)}
            {\(\mpcstate{\mach}{\pol}{\context} \stepstocon{\obs}
               \mpcstate{\mach'}{\pol'}{\context'}\)}

\subsection{Traces}
\label{sec:traces}

A trace over a type \(\TRACEELEMS\) is a nonempty, finite or infinite sequence of
\(\TRACEELEMS\)s; we write \(\TRACE{\TRACEELEMS}\) for the set of traces over
\(\TRACEELEMS\). We will mostly be concerned with sequences of MPC-states,
called \emph{machine traces} and denoted \(\MPCT \in \TRACE{\MPCS}\), and of observations.

\paragraph*{Trace-Of}

The ``trace-of'' operator, written \(\mpcstatename \hookrightarrow \MPCT\),
coinductively relates an initial state with the trace of states
produced by repeated application of step, and potentially cut short by
the policy component.

\judgment{\(\neg \exists \pol' . \mpcstate{\mach}{\pol} \polstep \pol'\)}
         {\(\mathit{obsforward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
           \mpcstate{\mach}{\pol}{\context},\obs\)}
%
\judgmenttwo{\(\mpcstate{\mach}{\pol}{\context} \stepstoobs[\obs'] \mpcstate{\mach'}{\pol'}{\context'}\)}
            {\(\mathit{obsforward} ~ \mpcstate{\mach'}{\pol'}{\context'} ~ \obs' = \machT\)}
            {\(\mathit{obsforward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
              (\mpcstate{\mach}{\pol}{\context},\obs) \machT\)}%
%
\judgment{\(\mathit{obsforward} ~ \mpcstatename ~ \tau = \machT\)}
         {\(\mpcstatename \hookrightarrow \machT\)}

\paragraph*{Projections}

We project out the machine state of an MPC-state with \(\pi_\mach\), the
policy state with \(\pi_\pol\), and the context with \(\pi_\context\). We overload
these operators to project out of an machine-trace a trace of machine states,
policy states, and contexts, respectively.

\paragraph*{Observations-Of}

The ``observations-of'' operator, written \(\pi_\obs\), takes a machine trace and
creates a trace of observations corresponding to the observations of each step in
the trace. It uses a coinductive helper \(\mathit{obsforward}\) that relates an
initial observation, a machine-trace, and an observation trace:

\judgmentthree{\(\pi_\mach ~ \mpcstatename \stepstoobs[\obs'] \pi_\mach ~ \mpcstatename'\)}
              {\(\mpcstatename' \hookrightarrow \MPCT\)}
              {\(\mathit{obsforward} ~ \obs' ~ \MPCT = \obsT\)}
              {\(\mathit{obsforward} ~ \obs ~ \mpcstatename \MPCT = \obs\obsT\)}%
\judgment{}
         {\(\mathit{obsforward} ~ \obs ~ \mpcstatename = \obs\)}

Then to take the observations from any particular trace, we start with a \(\tau\) observation:

\judgment{\(\mathit{obsforward} ~ \tau ~ \MPCT = \obsT\)}
         {\(\pi_\obs ~ \MPCT = \obsT\)}
\noindent
(We use juxtaposition of an element and a sequence to represent ``cons.'')

In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
to relate a machine-policy-context state triple to the MPCO-trace
induced by the step function.

\paragraph*{Head and Last}

We take the first element of a trace with \(\head(\Trace)\), which is a total
function since traces are non-empty, and the final element (if one exists) with
\(\last(\Trace)\), which is partial.

\paragraph*{Until}
The operation \(\PUT ~ f ~ \Trace\) takes a trace
\(\Trace \in \TRACE{\TRACEELEMS \times \OBSS}\) and a predicate
on elements \(f \subset \TRACEELEMS\) and gives the prefix of
\(\Trace\) ending with the first element on which \(f\) holds.

\begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f ~ \traceelem\)}
             {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
    \judgment{}
             {\(\PUT ~ f ~ \traceelem = \traceelem\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
    \judgmenttwo{\(\neg f ~ \traceelem\)}{\(\PUT ~ f ~ \Trace = \Trace'\)}
                {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\Trace'\)}
  \end{minipage}
\end{center}
%
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate always holds, then
\(\PUT\) is the identity function.

% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
%$\mathit{split}$: it takes two traces and links them by replacing the
%last element of the first (if it has one) with the first element of
%the second

%\judgment{} {\(t \cdot T = T\)}
%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
%\leo{Lemma/theorem: split/join = id?}
%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) is a predicate on machine states, we define \(\mpcstatename
\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
running \(\mpcstatename\) up to \(f\)'') and similarly for traces with policies:

\begin{center}
\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
\end{center}

\paragraph*{Context Segments}

In particular, we wish to formulate properties around consecutive subtraces
in which the context obeys a given predicate. This will commonly be used to extract
individual calls from a trace by taking subtraces in which the stack is
at or below a given depth. If \(f \subseteq \MACHS \times \CONTEXTS\), then
we have \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''

\begin{center}
  \judgmenttwo[Now]
              {\(f ~ \mach ~ \context\)}
              {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
              {\(\mach,\pol,\context \hookrightarrow \MPCT \langle f \rangle\)}
%
  \judgmentthree[Skip]
                {\(f ~ \mach ~ \context\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
%
  \judgmentthree[Search]
                {\(\neg f ~ \mach ~ \context\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT | f\)}
                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
\end{center}

\paragraph*{Reachability}

We often want to quantify over all traces from any initial state so that our
properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).

\paragraph*{Observational Similarity}

We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
as far as an external observer is concerned, written \(\obsT_1 \lesssim
\obsT_2\), if the sequence of
non-silent observations of $\obsT_1$ is a prefix of those of
$\obsT_2$; that is, we operate up to deletion of \(\tau\) observations,
coinductively:

\begin{minipage}{.3\textwidth}
  \judgment{}{\(\obsT \lesssim \obsT\)}
\end{minipage}
\begin{minipage}{.3\textwidth}
\judgment{}{\(\tau \lesssim \obsT\)}
\end{minipage}
\begin{minipage}{.3\textwidth}
\judgment{}{\(w \lesssim w\obsT\)}
\end{minipage}

\begin{minipage}{.3\textwidth}
\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(\tau \obsT_1 \lesssim \obsT_2\)}
\end{minipage}
\begin{minipage}{.3\textwidth}
\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(\obsT_1 \lesssim \tau \obsT_2\)}
\end{minipage}
\begin{minipage}{.3\textwidth}
\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(w\obsT_1 \lesssim w\obsT_2\)}
\end{minipage}

\smallskip
We then define similarity of observation traces as traces prefixing each other:
\[\obsT_1 \simeq \obsT_2 \triangleq \obsT_1 \lesssim \obsT_2 \land \obsT_2 \lesssim \obsT_1\]
%
%\leo{I still don't know how to do this transition. Maybe contrast
%with the ``perhaps surprisingly'' later?} Note that
%similarity of observation traces could alternatively be defined coinductively,
%just like the prefix relation, but by dropping the asymmetric rules that allow
%for a finite trace ($\tau \lesssim \obsT$ or $w \lesssim w\obsT$).\apt{Is this immportant?}\leo{I agree. Probably not}

Note that an infinite silent trace is a
prefix of (and similar to) any other trace. While this might seem
surprising at first, it makes sense in a timing-insensitive context:
an external observer looking at two machine runs cannot (computably)
distinguish between a machine that steps forever and a machine that
steps for a long time before producing some output.

\section{Local State Encapsulation, Formally}
\label{sec:lse}

We are finally ready for our definition of stack safety. We begin
by describing how we model knowledge about the program structure,
then dive into the formal definitions of local state encapsulation and
well-bracketed control flow, which together constitute stack safety.
Initially we will focus on a simple stack model, with a single stack
that doesn't share between caller and callee. All arguments are passed
in registers.

\subsection{Program Structure}

Our properties and policies assume that the program is annotated
by the compiler with information about its structure. At minimum,
we need to know which addresses contain instructions, which instructions
constitute calls and returns, and to which function an instruction
belongs.

This information is represented by a {\em code map}, a partial function from
addresses to pairs consisting of a function identifier and an {\em annotation}.
An annotation may be {\em call} or {\em return}.

\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap\})\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option{\ann}\]

\paragraph*{Call Depth and Domains}

At any given step of a program, some number of callers have pushed their frames
onto the stack, ``sealing'' that memory for later use; this number is the current
{\em call depth}. The call depth, as well as the association between stack addresses
and sealed frames, are dynamic and cannot be deduced from the state alone, so we will
track them over the course of an execution as additional context that informs the property.

Each component will be assigned to a {\em domain}, which may be {\em Outside} of the
stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed} by a particular
depth.

\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
  \times (\depth \in \mathbb{N}) \]

An initial context \(\context_0 = \domMap_0, 0\) where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
Then at each step we update the context according to the context-step function,
\(\stepsto_C\), which we will define in terms of the operations
\(\push{\domMap}{n}{\components}\), in which \(\components \subseteq \COMPONENTS\),
and \(\pop{\domMap}{n}\), pronounced ``push'' and ``pop'' and defined as follows.

\[
(\push{\domMap}{\depth}{\components}) ~ \component =
\begin{cases}
  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
                    \text{ and } \component \in \components \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]
\[
(\pop{\domMap}{\depth}) ~ \component =
\begin{cases}
  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1} \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]

Then we can define the context update relation. Let \(\sealcon\) be the sealing convention
that maps a machine state to a set of components, typically the set of addresses below
the stack pointer. The {\sc Default} rule applies only if neither of the others do.

\begin{center}
\begin{minipage}{.4\textwidth}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \callmap\)}
              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
\end{minipage}
\hspace*{0.1\textwidth}
\begin{minipage}{.4\textwidth}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
\end{minipage}
\end{center}

\begin{center}
  \judgment[Default]
           {\(\mach \stepsto \mach'\)}
           {\(\mach,\context \stepstocon \mach',\context\)}
\end{center}

\subsection{Ultra-eager Integrity}

Our first property will be {\em ultra eager}, meaning that it will be quantified over each step
of execution from any initial state. The criterion for integrity is simple: sealed components do not
change.

\definition A system enjoys {\em ultra-eager integrity} if for any reachable state
\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon{} \mpcstatename'\),
and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
\pi_\mach(\mpcstatename') ~ \component\).

\subsection{Eager Integrity and the Call Rule}

\definition Let \(\mpcstatename\) be a reachable state whose PC value is annotated
as a call and whose call depth is \(\depth\). Assume \(\mpcstatename \stepstocon \mpcstatename'\).
Let \(\MPCT\) be the trace ending after the corresponding return, if any, defined by
\(\mpcstatename' \hookrightarrow \MPCT | f\) with \(f ~ \mach ~ \context\) holding when
\(\context = \_,\depth\). Finally let \(\components\) be the set of components \(\component\)
such that \(\sealcon ~ \pi_\mach(\mpcstatename) ~ \component\) holds. Then by the {\em Call Rule},
it is sound to assume that if \(\last(\pi_\mach(\MPCT)) = \mach'\), then
\(\pi_\mach(\mpcstatename) ~ \component = \mach' ~ \component\).

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold when
\(\depth' \geq \depth\). A system enjoys {\em eager integrity} if each reachable segment
\(\MPCT \langle f_\depth \rangle\) satisfies \(\pi_m(\head(\MPCT)) ~ \component =
\pi_m(\last(\MPCT)) ~ \component\) for all \(\component\) where \(\domMap ~ \component =
\sealed{\_}\).

\subsection{Eager Confidentiality and the Insensitivity Rule}

\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if, for
all \(\component \not \in \components\), \(\mach ~ \component = \nach ~ \component\). We write
this \(\mach \approx_\components \nach\).

\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) have the {\em same differences}
if for all components \(\component\) such that \(\mach ~ \component \not = \mach' ~ \component\) or
\(\nach ~ \component \not = \nach' ~ \component\), \(\mach' ~ \component = \nach' ~ \component\).
This is written \(\mach,\mach' \diamond \nach,\nach'\).

\definition Let \(\mpcstatename\) be a reachable call state. Assume the that the call is successful
and \(\mpcstatename \stepstocon \mpcstatename'\). Then let \(\npcstatename'\) be a \(\components\)-variant
where \(\components\) is the set of all addresses in the stack. Then by the {\em Insensitivity Rule},
if the trace from \(\mpcstatename'\) eventually returns, so does that from \(\npcstatename'\),
their states after the return have the same differences, and up to return their traces are observationally
equivalent.

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold for
\(\depth' \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth),\_\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
Then take any \(\nach\) such that \(\mach \approx_\components \nach\). \(\MPCT\) corresponds to the
trace of a subroutine from its call to its return.

If we take the induced trace from this variant until a return,
\(\nach,\pol,\context \hookrightarrow \NPCT | \neg f_\depth\), we get a subtrace corresponding
to the same subrouting called with the same (register) arguments but with a potentially different
state at the call-site. A system enjoys {\em eager confidentiality} if for all such pairs of traces,
they correspond according to one of four cases:

\begin{enumerate}
\item \(\MPCT\) may be finite and \(\pi_\context(\last(\MPCT)) = (\_,\depth')\) with \(\depth' < \depth\),
  and \(\pi_\context(\last(\NPCT)) = (\_,\depth')\) with \(\depth' < \depth\). That is, both traces end
  having left the call of depth \(\depth\). Further:
  \begin{itemize}
  \item They have the same observable behavior, \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
  \item Their final states have the same differences, \(\pi_\mach(\head(\MPCT)), \pi_\mach(\last(\MPCT)) \diamond
  \pi_\mach(\head(\MPCT)), \pi_\mach(\last(\NPCT))\)
  \end{itemize}
\item \(\MPCT\) and \(\NPCT\) may both be infinite, with \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
\item \(\MPCT\) may be finite but end without returning -- which can only have happened if it fail-stopped --
  and its observable behavior prefixes that of \(\NPCT\), \(\pi_\obs(\MPCT) \lesssim \pi_\obs(\NPCT)\)
\item \(\NPCT\) may end with a fail-stop as in case 3, with \(\pi_\obs(\NPCT) \lesssim \pi_\obs(\MPCT)\)
\end{enumerate}

\subsection{Stack Safety with Sharing}
\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

\subsection{Trace Properties}

\subsection{Local State Encapsulation}

\definition A system enjoys {\em local state encapsulation} if it enjoys eager integrity and
eager confidentiality.

\definition A system enjoys {\em lazy local state encapsulation} if it enjoys lazy integrity
and eager confidentiality.

We do not give a confidentiality equivalent to lazy integrity, because the family of lazy
enforcement mechanisms that we consider under the HOPE system in fact enforce eager confidentiality.
A policy based on taint-tracking might require such a lazy confidentiality, and it would be constructed
by replacing the ``same-difference'' condition with one that requires any differences to produce
observationally equivalent states. But for our purposes it is irrelevant.

\section{Coroutine Safety}
\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

\subsection{Well-bracketed Control Flow}
\label{sec:wbcf}

Both the integrity and confidentiality components of local state
encapsulation concern themselves only with the accessibility of data,
and may still hold on a wide range of programs that violate expected
control flow. The archetypal example of this is the {\em awkward example},
in which an adversarial callee stores a return address and returns to it
on a subsequent call, thus violating the expected control flow of the
caller without actually violating the caller's integrity. But it also
encompasses simpler violations, such as a function bypassing a password
check by jumping (or returning) in the middle of a block of code.
%
Stack safety enforcement mechanisms like \citet{Skorstengaard+19}
usually aim to prevent such unwanted flows: every jump between
functions should either be a call or a return, and each return should
be to the instruction after its corresponding call.

To formally capture this aspect of stack safety, we introduce some more
security properties. {\em Control separation} restricts transitions between
functions to follow the interfaces set out by call, return, and yield, forming
a foundation for all of the other properties that rely on those interfaces for
meaning. {\em Entry integrity} ensures the validity of call targets,
{\em return integrity} ensures that of return targets, and
{\em yield-back integrity} ensures that of yield targets.

Recall that our properties use a compiler-derived code-map determine
which addresses are instructions and which functions they belong to, as well
as annotating them as calls, returns, and yields, as well as sharing information.
We now extend it with an annotation for valid entry points:
\[\ann \in \ANNS ::= \mathscr{P}(\{\callmap, \retmap, \yieldmap, \entmap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

\definition%{Control Separation}

Let \(\mpcstatename\) be a reachable state whose program counter points to code in
function \(\funid\), and \(\mpcstatename \stepstopol \mpcstatename'\) where
\(\mpcstatename'\) is state whose program counter points to a different function
\(\funid'\). A system enjoys {\em control separation} when all such \(\mpcstatename\)
are annotated \(\callmap\), \(\retmap\), or \(\yieldmap\).

\definition%{Entry Integrity}
A system enjoys {\em entry integrity} when, for any reachable \(\mpcstatename\) and
\(\mpcstatename'\) such that \(\mpcstatename\) is annotated \(\callmap\) and
\(\mpcstatename \stepstopol \mpcstatename'\), \(\mpcstatename'\) is annotated \(\entmap\).

\definition%{Return Integrity}
A system enjoys {\em return integrity}... [this one needs to talk about context.]

\definition%{Yield-back Integrity}
A system enjoys {\em yield-back integrity}... [same.]

\definition% {Well-Bracketed Control Flow}
The above properties are somewhat orthogonal, but together they
capture a standard notion of control flow for a stack: functions keep
their code separate outside of calls and returns, calls respect
explicit entry points, and returns respect implicit entry points
corresponding to their call.  A system with all three
properties (with respect to a function map, call map, return map,
and entry map) is said to enjoy
{\em well-bracketed control flow} (with respect to those maps).

\section{Enforcement}
\label{sec:enforcement}

In this section we examine how an existing enforcement mechanism,
Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, can implement
the formal stack-safety property described in the previous section: it in
fact enforces a stronger, inductive version of that property which strengthens the
local state encapsulation component by checking for integrity and
confidentiality violations at every step of the callee's execution trace.

%\paragraph*{Micropolicies}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program.  To control this monitor, they use a programming model that
allows fine-grained manipulation of metadata tags to encode so-called
\emph{micropolicies}~\citep{pump_oakland2015}.
In such a system, all values in
memory addresses and registers (including the $\PCname$)
are enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micropolicy is defined as a set of tags and a
collection of software-defined rules, indexed by machine opcode.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs (PC, registers, memory) and on the instruction itself, and produces
one of two outcomes: either the instruction is allowed to execute (and generates
tags for the result of the operation and the new PC), or the machine
fail-stops with a policy violation.
\citeauthor{pump_oakland2015} have shown that a wide range of micropolicies can
be defined using this scheme.

Efficient execution of these micropolicies relies on hardware implementations,
such as the PUMP architecture~\citep{pump:asplos2015}.  Tags are represented
as word-size bit vectors, stored separately from the regular memory and registers.
The hardware incorporates a rule cache
to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
the hardware traps to a software handler (running in a privileged context or
on a co-processor) to compute the rule result. To obtain adequate performance,
it is important to design micropolicies so that they hit in the cache as
much as possible. Thus, practical policies maintain a small working set of
distinct tags.
%% %
%% \rb{Maybe go into more detail, examples later\ldots or try to segue into stack
%% policies}

\subsection{A Conservative Tag Policy}
\label{sec:conservative}
%
The first micropolicy we present to enforce the definition of stack
safety developed in \cref{sec:lse-and-wbcf} is a small variation
on the Depth Isolation policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

Our micropolicy works as follows. (We assume for ease of exposition that no
arguments are passed on the stack.)
%% \rb{Come back to this later?}.
The micropolicy maintains tags on values in memory of the form $\tagStackDepth{n}$,
indicating locations that belong to the stack frame at activation depth $n$,
or $\tagNoDepth$; it tags the $\PCname$ with tags of the form $\tagPCDepth{n}$.
During normal execution, the micropolicy rules only permit load and store operations
when the target memory is tagged with the same depth as the current {\PCname} tag.
Initially, the entire stack is tagged $\tagNoDepth$,
%\rb{actually, the ``initial frame'' should be tagged with its proper depth}
and the {\PCname} has tag $\tagPCDepth{0}$. These tags are altered
at exactly those points in
the program where the contour of the stack-safety property changes:

\begin{itemize}

\item From caller to callee, when the machine executes an instruction marked as a call in
  the call map.
  At this point, the current
  $\tagPCDepth{n}$ tag is incremented to $\tagPCDepth{(n + 1)}$, and the function
  entry sequence initializes  all locations of the new function frame with
  tag $\tagStackDepth{(n + 1)}$.

\item From callee back to caller, when execution is about to reach the return point
  $\ret{c}$
  corresponding to the call at $c$.  The exit sequence retags all locations in the function
  frame with $\tagNoDepth$. At the actual return instruction, the
  $\PCname$ tag is decremented.

\end{itemize}

This discipline suffices to enforce local stack encapsulation.
To implement it,  we define ``blessed'' instruction sequences
intended to appear at the entry and exit of each function,
which manipulate tags as just described in addition to performing the
usual calling convention tasks of saving/restoring the return address to/from
the stack and adjusting the stack pointer. With the aid of an additional tag on
the $\rsp$ register, these sequences also serve to enforce
well-bracketed control flow.
The micropolicy guarantees atomic execution of these sequences
using a combination of tags on the instructions
and an additional tag on the $\PCname$; we omit the details here.

There remains the question of how to ensure that the sequences are
invoked at the right places. For a program to enjoy the stack-safety property
(with respect to a particular call map), all we need require is that
the entry sequence be initiated at any instruction marked as a call
in the call map. This is achieved by giving these instructions (another) special tag,
and it is easy to check statically that this has been done correctly.
If the code fails to initiate an exit
sequence at a point where the stack-safety property expects a return,
the micropolicy will incorrectly behave as if execution
is continuing in the callee, but since the callee never has stronger access
rights than the caller, this is harmless.

%% APT: removed because the tag part is too mysterious and the instruction part
%% is standard.
%% This is the entry
%% sequence:
%% %
%% \setcounter{pcctr}{1}
%% \[
%%   \begin{array}{l|l|l|l}
%%      & \mathit{Instruction} & \mathit{Tags} & \\
%%     \hline
%%     \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
%%         {Store return address in stack}
%%     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
%%         {Increment stack pointer by frame size}
%%   \end{array}
%% \]
%% %
%% And the exit sequence:
%% %
%% \setcounter{pcctr}{1}
%% \[
%%   \begin{array}{l|l|l|l}
%%      & \mathit{Instruction} & \mathit{Tags} & \\
%%     \hline
%%     \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
%%         {Load return address}
%%     \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
%%         {Decrement stack pointer by frame size}
%%     \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
%%         {Return to caller ($\jalr$)}
%%   \end{array}
%% \]

%% \rb{Positioning of the above sequences, relation to running example. The tagging
%%   of stack frames would be considered part of these sequences.}

%% These sequences also enforce the $\SP$ discipline. When stack frames
%% are of fixed size, it suffices to increment $\SP$ and tag it with a
%% dedicated tag in the header sequence, which is later checked after
%% decrementing $\SP$ during the exit sequence to authorize the
%% return. The well-formedness of the blessed sequences can be checked
%% statically.

%% Only programs whose sequences are well-formed are
%% protected by the micropolicy, although programs are still protected
%% (by fail-stopping) even if the intended sequences are missing. By
%% tracking the execution of the blessed sequences in the tag state of
%% the {\PCname} register, the micropolicy ensures that the entry
%% sequences of calls declared in the call map are executed correctly.

Note that in this policy both the entry and exit sequences
must write to each element of
the frame, which can be quite expensive, especially for programs that
allocate large but sparsely populated frames; we return to this point
in \cref{sec:lazy}.
%
The running example in \cref{fig:running-trace-a} manages the stack in this way.
For example, {\tt f}'s ``blessed'' sequences allocate space to save its return
address (and private variables) in instructions 20--21, and later restores that
return address and deallocates its private frame space before returning in
34--36. Additionally, unlike in the simplified presentation of the policy, the
caller passes information to the callee through the stack, and for this creates
and initializes (22 and 24--25) and destructs (29) the argument part of the
callee's stack frame.\rb{The policy would also need to control this, I think it
  is clear}
%
%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}

%% Because the access rights of callers subsume those of callees, it is essential
%% that the micropolicy change tags when the property thinks that a call occurs,
%% but it is safe for the micropolicy to fail to reset to the caller's tag when the
%% property thinks a return occurs. (Dually, it is safe for the micropolicy to
%% change tags even without executing a call point, but not for it to reset tags to
%% the caller when the property thinks execution is still in the caller).

%% Because the property does not identify return sites, we have no way to require
%% that ``intended returns'' are tagged correspondingly. Some unintended results
%% might also be so tagged, but as long as the jump has the effect of a legal
%% return, it is OK.

\section{Extension to Cheri}

\section{Extension to Software-based Enforcement}

\section{Stack Safety Properties for Lazy Enforcement}
\label{sec:lazy}

The conservative policy described in the previous section, while
testable and enforceable, is rather slow to be of practical use. What
we want, instead, is a somewhat more permissive policy that can also
be efficiently implemented. Most of the performance overhead incurred
stems from the need to set stack activation tags as a frame is created
and reset them as it is destructed, as
\citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
mitigate those costs, they propose optimizations called Lazy Tagging and
Lazy Clearing that relax and defer some of the checks of the Depth Isolation
policy. Under these optimizations, the policy does not initialize stack
frames on entry or clear them on exit (thus speeding up execution),
and it permits all writes to the stack, even when the $\PCname$ tag
does not match the memory tag.
However, reads from the stack do still require $\PCname$ tag and memory tag
to match. Thus, even if a callee illicitly writes to a private location
in its caller, the caller will eventually detect this if it ever tries to read from
that location.
%
These \emph{lazy policies} admit more efficient implementations, but
they deliberately allow violations of stack integrity temporarily,
with the checks deferred until the point a violation truly becomes
harmful. The natural question then is, how do we characterize the
protections provided by these policies once stack safety is broken?
How does a harmful violation that must be caught later differ from a
harmless one? The properties in this section rely on our notion of
observations to draw this distinction.

\subsection{Observable Integrity}

\paragraph{Rolling Back to Idealized State}

Observable integrity weakens eager integrity by allowing a returned
state to differ from its call state at high integrity components,
provided that the differences do not change the observable behavior of
the rest of the program. The challenge here is to define what constitutes
a change in behavior. We construct an idealized return state that
contains only the changes that the caller was allowed to make, and use
the trace from that state as a baseline against which the real trace
will be compared.

      \definition

      Given a contour, a call state, and a returned state, we define a
      {\em rollback} function \(\mathit{roll} \in \CONTOURS \times \MACHS \times
      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
      the call state on those components that are high integrity in its
      contour, and matches the return state on low integrity components:
      \[\mathit{roll}(\contour,\mach_e,\mach_r)(\component) =
      \begin{cases}
        \mach_e(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
        \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
      \end{cases}\]

      If \(\mach_e\) is an entry to a call with contour \(\contour\)
      and \(\mach_r\) its return, the result of
      \(\mathit{roll}(\contour,\mach_e,\mach_r)\) agrees with the
      entry \(\mach_e\) on every component the callee should not
      change, and with the actual return \(\mach_r\) on those
      components that it is permitted to change.  So the observable
      behavior that follows is the idealized behavior to which we
      compare the trace from \(\mach_r\).

      Now we can define a version of trace integrity that compares the trace
      from a returned state to that from its idealized counterpart.

    \definition%{Observable Trace Integrity}

      A trace enjoys {\em observable trace integrity} with respect to
      some contour \(\contour\) if the trace induced by its final
      state (if any) has observable behavior that prefixes that of its
      idealized version (to account for premature halts due to policy
      violations):

      \[\begin{split}
        \propdef{ObsTraceInt} ~ \contour ~ \MPT \ \ \triangleq \ \
        & \head(\MPT) = \mpostate{\mach_e}{\_}{\_} ~ \land ~
          \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
          \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
          \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

    \begin{figure}
      \integritylazyexample
      \caption{Rolling back from Figure \ref{fig:intex}.}
        %\apt{But at 27, $[\rsp]$ contains 0, not 18, because of the change at 44. Could just fix diagram, but this complicates story.}}
      \label{fig:intlex}
    \end{figure}

    For our example, we return to the violation of trace integrity in
    \cref{fig:intex}, in which {\tt g} has written to {\tt z} and returned.
    Whether observable integrity holds depends on the continuing trace, shown
    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
    \(\mach_r\), and the rolled-back state, \(\mach_r' =
    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
    components are restored to their values as of \(\mach_e\). Specifically,
    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
    it prints 0. Only then is observable trace integrity violated.

    However, consider instead if we modify the original code of \cref{fig:running-trace-b} to insert after instruction 44 a new
    instruction \(44.5: [\rsp-4] \leftarrow 50\), overwriting {\tt w} in
    {\tt f}'s frame. So \(\mach_r'\) maps address 105 to 50, while the
    rolled-back state \(\mach_r'\) maps it to 0. But in the trace from either
    state, instructions 27 and 28 load the return value (18) and store it to
    105, eliminating the only difference. The remainder of the observation
    traces must be identical, so this trace does obey observable trace
    integrity.

    \definition%{Observable Stack Integrity}

      A system enjoys {\em observable stack integrity}, like stack integrity,
      if, for any initial state and each \(n\)-transition from
      \(\mpstate{\mach_c}{\pol_c}\) to \(\mpstate{\mach_e}{\pol_e}\) in its
      induced trace, we take the prefix of the call
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\)
      and observable trace integrity holds on \(\MPT_{\mathit{call}}\) with
      contour \(\mathit{Cof}(\mach_c,n)\).

  \subsection{Observable Confidentiality}

      Two of the cases of trace confidentiality (\cref{def:trace-conf}), \(\mathit{InfCase}\)
      and \(\mathit{FailCase}\), are defined solely in terms of
      observation traces. If
      the traces are infinite, they should be similar; if one
      fail-stops, it should be a prefix. Only the
      \(\mathit{ReturnCase}\) checks machine states, because in that
      case the traces could continue, but have been cut off by a
      return. The key to making confidentiality lazy is not to perform that
      check but rather to consider the traces after that return,
      checking only whether they are observationally similar.

      This requires restoring varied components to their original
      values in the trace from the variant state, since after return
      they will once again be legitimately accessible. To return to
      the example in \cref{fig:confex}, the varied final state
      \(\mach_r'\) still has $-5$ as a return address for {\tt f}, so
      the trace after that return will behave oddly in a way that
      does not reflect a confidentiality violation.  That means we
      cannot immediately compare traces from \(\mach_r\) and
      \(\mach_r'\), but rather that we must restore the varied values
      while preserving differences that have occurred during
      execution.

      \definition

      The \(\mathit{restore}\) function takes four states: the
      original entry state, a variant entry state, and corresponding return
      states. It returns another state. The resulting state restores those
      components that varied between entries to match the original entry,
      unless they were changed during execution:
      \[\mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')
        (\component) =
        \begin{cases}
          \mach_e(\component) & \text{if } \mach_e(\component) \not =
                                \mach_e'(\component) \text { and } \\
          & \mach_e(\component) = \mach_r(\component) \text{ and }
            \mach_e'(\component) = \mach_r'(\component) \\
          \mach_r'(\component) & \text{else} \\
        \end{cases}\]

      Now we can describe a trace property akin to trace confidentiality, but
      continuing execution after the end of the callee trace.

    \definition{%Observable Trace Confidentiality}

      {\em Observable trace confidentiality} is the lazy analog to
      trace confidentiality, and reuses the cases in which the traces
      do not return.  For a return predicate \(R\) and a pair of
      traces \(\MPT\) and \(\machT\), we consider the return case, in
      which $\MPT$ has a last element
      \(\mpostate{\mach_r}{\_}{\_}\) and that machine state is
      a return, i.e. \(R ~ \mach_r\) holds.
%
      Observable trace confidentiality then requires that $\machT$
      also has a final state \((\mach_r',\_)\) that is a return, i.e.
      \(R ~ \mach_r'\) holds.
%
      In addition, just like in the eager case,
      observations of \(\MPT\) are similar to those of \(\machT\).

      The only thing that changes is the requirement that \(\mach_r\)
      and \(\mach_r'\) agree on components. Instead, let \(\mach_e\)
      be the machine state of \(\head(\MPT)\), and \(\mach_e'\) be
      that of \(\head(\machT)\). We restore \(\mach_r'\) based on
      \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).  Then
      observable trace confidentiality requires that the
      induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
      observational prefix of the induced trace from \(\mach_r'\),
      reflecting the possibility of a fail-stop after return.

      \[\begin{split}
        \mathit{ObsReturnCase} ~ R ~
        (\mpostate{\mach_e}{\_}{\_}\MPT) ~
        ((\mach_e',\_)\machT) \ \ \triangleq \ \
        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land ~
         \last(\machT) = (\mach_r',\_) ~ \land \\
        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') ~ \land \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
        \mach_r'' \hookrightarrow \machT' \Rightarrow \\
        & \pi_o(\MPT) \eqsim \pi_o(\machT) ~ \land ~
        \pi_o(\MPT') \lesssim \pi_o(\machT') \\
      \end{split}\]

      We combine this case with the other two cases from ordinary trace
      confidentiality:
\[
        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \ \ \triangleq \ \
         \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT ~ \land
         \mathit{InfCase} ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
          \MPT ~ \machT
\]

      \begin{figure}
        \confidentialitylazyexample
        \caption{Continuing from \cref{fig:confex}.}
        \label{fig:conflex}
      \end{figure}

      In our example, \cref{fig:conflex} shows how execution continues from
      \(\mach_r'\) and the restored
      \[\mach_r'' = \mathit{restore}(%\contour,
      \mach_e,\mach_e',\mach_r,\mach_r').
      \]
      Addresses 106 and 107
      are not restored; every address that was varied and not changed in one
      of the executions is. So 107 contains 43 in the trace from \(\mach_r\),
      and after returning from {\tt f}, {\tt main} prints 85. But in the trace
      from \(\mach_r''\), 107 contains $-4$, and ultimately 38 is printed.
      So observable trace confidentiality does not hold. This example
      illustrates the weakness of observable confidentiality as a testing
      property: it takes until the very end of the program to determine whether
      a tainted value actually influences observable behavior.

      On the other hand, suppose that we inserted an instruction at 43:
      \(r_6 \leftarrow [\rsp-8]\). \(r_6\) would hold 42 in \(\mach_r\)
      and $-5$ in both \(\mach_r'\) and \(\mach_r''\). But since \(r_6\) is
      never used otherwise, the traces of {\tt f} will enjoy observable trace
      confidentiality.

    \definition%{Observable Stack Confidentiality}

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. For each initial
      state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
      \(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\), take any
      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~
      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} ~ | ~
      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
      \ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}\).

      Additionally, for any variation \(\mach_0' \approx_{\mathit{Cof}(\mach_0),0}
      \mach_0\), taking the induced trace  \(\mach_0' \hookrightarrow
      \machT\), we must have \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\).

      A system enjoys observable stack confidentiality if observable
      confidentiality holds in all such cases.

    \definition%{Observable Stack Safety}
    A system enjoys {\em observable stack safety}
    % with respect to a call map       and a function map
    if it enjoys both observable stack integrity and
      observable stack confidentiality with respect to them.

      \medskip
      \textsc{Theorem.}
      Stack safety implies observable stack safety.
      \smallskip

      This is more straightforward than the implication in the previous
      section (1,5K LoC in Coq) as the recursive
      structure of the two properties is similar.

    \paragraph{Connection to Policies}

    Observable properties allow us to defer enforcement until a property
    violation would become visible, as in lazy policies. But by formalizing
    the connection to the eager policy, we can identify when lazy policies
    miss violations that can become visible. The optimized lazy tag policy from
    \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
    at which it was written, and enforces that it must be read from the
    same depth. However, this allows violations of observable stack safety!
    Say we have functions {\tt foo}, {\tt bar} and {\tt baz}, where {\tt foo}
    calls {\tt bar}, and then after {\tt bar} returns {\tt foo} calls
    {\tt baz}. Since {\tt bar} and {\tt baz} are at the same stack depth, if
    {\tt bar} writes to {\tt foo}'s frame, {\tt baz} can then read the tainted
    write and use it to create an observable difference that violates both
    observable integrity and confidentiality without violating the policy.

    The lazy policy can be repaired if, instead of tagging each function
    activation with its depth in the stack, we generate a fresh activation
    identifier on each call, which prevents a program from exploiting stale
    tags from previous activations. A related mechanism was explored in the
    Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
    associates a unique activation identifier to each function, which is
    however shared by all activations of the same function in order to obtain
    good cacheability of tag rules. Static Authorities comes closer to
    observable stack safety, though the example above would still exist
    if {\tt bar} and {\tt baz} were merged into a single function that tracked
    whether it had been called previously and changed its behavior accordingly.
    \ifaftersubmission\apt{More blatently, static authorities doesn't work for recursive calls.}\fi

\section{Related Work}
\label{sec:relwork}

%% CHERI \rb{cite}

\paragraph{Formal Stack Safety on Capability Machines}
%
To our knowledge, the only line of work to date that has attempted
a positive and formal characterization of
stack safety is \citet{Skorstengaard+19b}, who introduce a calling
convention that uses local capabilities to preserve local state
encapsulation and well-bracketed control flow, using a logical
relation to reason about the stack safety of concrete
programs. Although the required hardware support is readily available
in capability machines like CHERI\bcp{including linear capabilities?? or
  don't they use them in this paper?}, this technique incurs significant
costs\bcp{This is the pot calling the kettle black}, because it requires the entire unused part of the stack to be
cleared whenever a security boundary is crossed. Their logical
relation captures capability safety without ``externally observable
side-effects (like console output or memory access traces).'' \bcp{Not very clear:}In the
discussion, ``while [the authors] claim that [their] calling
convention enforces control-flow correctness, [they] do not prove a
general theorem that shows this, because it is not clear what such a
theorem should look like,'' noting that the correctness property
enjoyed by their technique ``is not made very explicit.''

StkTokens \citep{Skorstengaard+19} continues this line of work. Like
the earlier paper, it aims to protect the stack by enforcing local state
encapsulation and well-bracketed control flow. It does so by defining a new
calling convention that makes use of linear capabilities for stack and return
pointers. The convention operates on a single shared stack and requires that
protected components avoid compromising their own security by following certain
simple rules---like not leaking their private capability seals. More
precisely but still informally, local state encapsulation is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow as only allowing returns from the topmost
frame to the immediately adjacent frame below. Formally, it improves on
\citet{Skorstengaard+19b} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return instructions,
which is proven fully abstract with respect to a more concrete capability
machine that replaces those pseudo-instructions with their calling convention.
Their proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces; their model of
observations is limited to cotermination.

We believe that the linear capability machines introduced by
\citep{Skorstengaard+19}, in combination with the StkTokens calling convention,
can be modeled in our framework and satisfy our definition of stack safety in
\cref{sec:lse-and-wbcf}. Further, we expect them to satisfy the even stronger
property developed in \cref{sec:enforcement}. Proving these conjectures would
involve materializing their definitions of local state encapsulation and
well-bracketed control flow, which exist only in enmeshed form as part of the
semantics of the capability machines. The main practical limitation of StkTokens
is its reliance on linear capabilities, as it is unclear now they could be added
to practical capability machines, especially in terms of efficiency---previous
work on micropolicies \citep{yannis-report} has shown how to use that framework
to implement linear return capabilities.\bcp{But IIUC the Dover realization
  of micropolicies can't do this...?}

\paragraph{Protecting the Stack with Micropolicies}
%
\citet{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
attacks against stack data are in scope (but not side channels or hardware
attacks), and study the protection of stack data through three families of
micropolicies that tag stack objects with a pair of frame and object identifiers
used to validate accesses to the stack: Return Address Protection (which
prevents an adversary from overwriting designated return addresses), Static
Authorities (which only allows the code of a function to access the stack frames
of its own dynamic instances), and Depth Isolation (described in
\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
temporal locality of stack memory and local call graphs, as well as information
generated by the compilation toolchain, to strike various balances between
precision and cacheability, evaluated through benchmarks that demonstrate
limited performance overhead. In addition to the baseline ``eager'' policies,
they propose a number of lazy optimizations, notably Lazy Tagging and Lazy
Clearing, discussed in \cref{sec:lazy}, with associated improvements in
performance.

\paragraph{Heap Safety as a Security Property}

Heap safety, like stack safety, may be framed as a security property
in the form of noninterference~\citep{DBLP:conf/post/AmorimHP18}. Just
like~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
characterization of the meaning of (heap) memory safety, in this paper
we aim to do the same for the stack.\bcp{Let's describe their work in a bit
  more detail.}
%
Their model describing safety of shared stack-allocated
objects extends naturally to heap safety, complicated by the fact that
these objects are not deallocated by returns, and therefore a
function's privilege may increase or decrease after its entry
point. For instance, if a callee allocates a heap object and returns
the pointer to its caller, that object's addresses become accessible
in contradiction to a contour computed at the caller's entry. So such
an extension must enforce trace confidentiality and integrity
properties separately on continuous segments of each call, with
contours computed at each crossing between caller and callee.

\section{Future Work}
  \label{sec:future}

  We have presented a pure notion of stack safety that omits many complicating
  factors common to real systems. In particular, our model as presented exploits
  the simplifying assumption that privilege never increases during a function
  call, and that therefore a contour computed once on the entry to a function
  describes its privilege throughout. Our future plans are to extend the model
  with common language features, some of which violate this assumption. Here we
  describe the basic principles by which our model can be extended to some of
  these cases. We also plan to develop a testing framework for quickly
  checking the validity of different enforcement mechanisms.

  \paragraph{Stack-derived Pointers}

    Common programming idioms involve a caller passing a pointer to
    its local data into a callee. Using standard notions of pointer
    provenance~\citep{provenance}, we can extend our model to
    distinguish safe use of this idiom from true stack safety
    violations. We require additional annotation to identify which
    addresses within a function's stack frame correspond to distinct
    objects. Then a valid pointer to an object is one derived through
    legal arithmetic from its base address, and at each call point, an
    object's addresses are marked low confidentiality and integrity if
    a valid pointer exists in a register or in the transitive closure
    of the callee's accessible memory. All pointers to an object cease
    to be valid when it is deallocated along with its stack frame.

  \paragraph{Non-stack control flow}

    Control structures beyond well-bracketed calls and returns require
    modifications to the model. Tail calls, for instance, reuse the caller's
    stack frame for its callee, and every nested tail call returns
    simultaneously to the top non-tail-calling function. Under normal stack
    integrity, the first function in a chain of tail calls returns and violates
    integrity, while subsequent tail calls never actually return at all.
    Instead we need a variant that treats the entry of a tail call as the
    return of its caller.

    More complex is the addition of a coroutine model, in which multiple stacks
    respect stack safety internally, but might also yield to one another.
    Here we must distinguish three levels of integrity: accessible, in the same
    stack, and in another stack. Stack integrity requires that same-stack data
    be unchanged when a function returns, but other-stack data may change. A
    similar {\em yield integrity} property requires that, from a yield out of
    a coroutine to the next yield back into it, all accessible and same-stack
    data are unchanged. Confidentiality properties are split similarly.

  \paragraph{Random Testing}

  Part of the motivation for the strong lockstep stack-safety property
  is that it uncovers errors quickly, and therefore should be quicker
  to test, just like the smarter testing properties
  of~\citet{TestingNI:ICFP}. We have already developed a testing
  framework for debugging our policy enforcement for this property, and
  it would be interesting to explore whether a testing-amenable variant
  of the lazy property also exists.

% \rb{Notes from meeting:
%   \begin{itemize}
%   \item Explain how the theory applies to production compilers like GCC and
%     LLVM, which aspects are directly applicable, what are the effects of
%     alternative calling conventions. Other less charted waters: tail calls,
%     coroutines, jumps into supervisor mode.
%   \item CHERI and its operation, documentation and practical configurations and
%     operation.
%   \item Compare and contrast with StkTokens.
%   \item Under what circumstances can stack safety exist without an enforcement
%     mechanism?
%   \item Zeroing out of allocated space inside blessed sequences, eager vs lazy.
%     Possibility of blessed alloc and dealloc sequences, which also manipulate SP
%     and would not be legal without blessing. In such an extended setting, the
%     return sequence would check for matching addresses (it would be nice to have
%     testing for this!).
%   \end{itemize}
% }

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
\bibliography{bcp.bib,local.bib}

%% Appendix
%\appendix
%\section{Appendix}
%Text of appendix \ldots

\end{document}