### 2021-03-24 tex/props.tex
Merged Leo's risc-v work with my most recent and organized the directory

\documentclass[conference]{IEEEtran}
\input{macros}

\usepackage{geometry}
\usepackage{amssymb,amsmath}

\begin{document}

  \section{Preliminaries}

  \leo{Low level presentation point that slightly bugs me: }

    Values and addresses are words:
    \[\word,\addr \in \WORDS\]

    Components are words (addresses) and registers, including special registers:
    \[\reg \in \REGS\]
    \[\component \in \COMPONENTS = \WORDS + \REGS + \SP + \PCname\]

    A machine state maps components to values:
    \[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]

    Our properties are defined in terms of observations, which may be silent or a word of data:
    \[\obs \in \OBSS = \tau ~ | ~ \word \in \WORDS\]

    Execution is by\leo{is by?} a step function \(\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow
    \OBSS \times \MACHS\).

    \paragraph*{Policy}

    A policy is a general model of an enforcement mechanism,
    consisting of a set of policy states, \(\pol \in \POLS\) and a
    policy step function \((\mach, \pol) \stepstopol \pol' : \MACHS
    \times \POLS \rightharpoonup \POLS\).

    The initializer and step function are both partial functions,
    the former to reflect static guarantees (some initial states may not be valid at all) and
    the latter to reflect dynamic fail-stop behavior. We then define machine-policy pairs and
    a step function on them:

    \[\MPS = \MACHS \times \POLS\]
    \[\stepstoobs{}_P \subseteq \MPS \times \MPS\]

    \[\frac{\mach_1 \stepstoobs{\obs} \mach_2 \hspace{.5cm} (\mach_1, \pol_1) \stepstopol \pol_2}
           {\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}}\]

  \paragraph{Traces}

    A trace is a potentially infinite, nonempty sequence. We define a few operations on traces generically.
    The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\) and a predicate on elements \(f\) and
    separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:
    \[\frac{f\ t}
           {\mathit{split}(f,tT) = t, tT}\]
    \[\frac{\neg f\ t \hspace{.5cm} \mathit{split}(f,T) = T_1,T_2}
           {\mathit{split}(f,tT) = tT_1,T_2}\]

    The {\it join} operation \(T_1 \cdot T_2\) takes two traces and links them by replacing the last element
    of the first (if it has one) with the first of the second:
    \[\frac{}
           {T_1t \cdot T_2 = T_1 T_2}\]

    Note that both operations are defined coinductively. We identify infinite traces with \(\isinf(T)\).
    We define a ``trace-of'' operator \(\hookrightarrow\) that coinductively relates a machine state with the
    trace of machine states paired with observations produced by repeated application of step:

    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \mach_1 \hookrightarrow \machT}
           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT}\]

    Since the step function is total, if \(\mach \hookrightarrow \machT\), then \(\machT\) must be infinite.
    We will also wish to take a prefix of the trace from an initial state up through the first state where
    some condition holds, using {\it split}. Let \(f\) be a predicate on machine states. Then we define
    \(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
    \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
           {\mach \hookrightarrow \machT_1 | f}\]

    %\[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(f,\machT) = \machT_1,\machT_2}
    %       {\mach \hookrightarrow f | \machT_2}\]

    We define \(\hookrightarrow_P\) to relate a machine-policy pair to the trace of MP pairs and observations
    created by \(\stepstoobs{}_P\):

    \[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
           {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
              \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
           {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]
    \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
           {\mpstatename \hookrightarrow \MPT_1 | f}\]
    %\[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm} \mathit{split}(f,\MPT) = \MPT_1,\MPT_2}
    %       {\mpstatename \hookrightarrow f | \MPT_2}\]

    We project the machine states from either machine trace with \(\pi_m\) and the observations with \(\pi_o\).
    An observational prefix \(\obsT_1 \lesssim \obsT_2\) means that the trace \(\obsT_1\) of observations is
    a prefix of \(\obsT_2\) up to the deletion of \(\tau\) observations (once again, coinductively):

    \[\frac{}
           {\obsT \lesssim \obsT}\]

    \[\frac{}
           {\tau \lesssim \obsT}\]

    \[\frac{\obsT_1 = \tau \obsT_1' \hspace{.5cm} \obsT_1' \lesssim \obsT_2}
           {\obsT_1 \lesssim \obsT_2}\]

    \[\frac{\obsT_2 = \tau \obsT_2' \hspace{.5cm} \obsT_1 \lesssim \obsT_2'}
           {\obsT_1 \lesssim \obsT_2}\]

    \[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2' \hspace{.5cm} \obsT_1' \lesssim \obsT_2'}
           {\obsT_1 \lesssim \obsT_2}\]

    Equivalence of observation traces is defined as traces prefixing each other:
    \[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]

    We define policy-sensitive observational equivalence, which relates a machine trace and
    an {\it mptrace} whose observable behavior match it, up to a possible halt due to policy failure.

    \[\frac{\neg \isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \lesssim \mathit{Obs}(\machT)}
           {\MPT \sim_P \machT}\]

    \[\frac{\isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT)}
           {\MPT \sim_P \machT}\]

  \section{Observable and Eager Stack Safety}

    \paragraph*{Call Maps and Returns}

    A {\it call map} identifies machine states that represent a call, abstracting away the explicit
    calling convention. Formally it maps some machine states to a number of arguments:
    \[\callmap \in \CALLMAPS = \MACHS \rightharpoonup \mathbb{N}\]

    As it pertains to a typical calling convention, for purposes of understanding the properties,
    a call map may be considered to identify the state just before the \(\JAL\), in which space for
    arguments is already allocated. But the mechanism is more general.

    A state on which a call map is defined is referred to as a {\it call state}, and we consider
    the call to have returned the first time control returns to the instruction following the
    \(\JAL\) with the stack restored. Formally this is a relation between the states:

    \[\begin{split}
      \ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) = \next(\mach_c(\PCname)) \land \\
                                        & \mach_r(\SP) = \mach_c(\SP)
    \end{split}\]

    \leo{TODO: Initial state depends on call map}

    \paragraph*{Stack safety}

    A stack safety property is a property of a transition system, a policy, and a callmap
    that is decomposed into an integrity property and a confidentiality property.
    We will present eager and observational variants of stack safety.

    {\it Eager stack integrity} states that if a function call returns, the caller's stack data
    will be identical to what it was before the call. This implies the weaker {\it observable
    stack integrity}, which states that any changes to the caller's frame will produce the same
    observable behavior.

    \sna{Attempt 2...}
    {\it Observable stack confidentiality} states that data outside a callee's stack frame are secret,
    and may not interfere with the observable behavior of the program. That is, the callee's behavior must
    be identical regardless of the configuration of memory outside its frame, and it must not change the
    program state based on that memory in a way that causes future changes in behavior.
    {\it Eager stack confidentiality} additionally states that changes the callee makes to the program state
    must be identical regardless of the configuration of memory.

    Our top level properties are {\it eager stack safety} and {\it observable stack safety}:
    \[\begin{split}
      & \textit{observable stack safety} \triangleq \\
      & \textit{observable stack integrity} \land \\
      & \textit{observable stack confidentiality} \\
    \end{split}\]
    \[\begin{split}
      & \textit{eager stack safety} \triangleq \\
      & \textit{eager stack integrity} \land \\
      & \textit{eager stack confidentiality} \\
    \end{split}\]

    We will separately introduce confidentiality at the top level, to illustrate it simply before
    moving into calls. Integrity at the top level is vacuous because there is nothing to protect.

  %  We unify these properties with a structure
  %  called a {\it contour} that maps components to levels of confidentiality and integrity.
  %  Specifically, we have confidentiality labels \(\mathit{labelC} ::= \{\HIGHSEC, \LOWSEC\}\)
  %  and integrity labels \(\mathit{labelI} ::= \{\HIGHINT, \LOWINT\}\), and a
  %  type \(\mathit{contour}\)

    \paragraph{Top-level Confidentiality}

    \leo{Do we really need safe initialization? Couldn't we just use the definitions of
      integrity and confidentiality for the toplevel, pointing out that since it never
      returns only confidentiality is interesting? Or would we miss something?}
    \sna{Changed to make it confidentiality, but still introducing contours after. Maybe
         contours should be before?}

      In the initial state of a program, nothing has yet been allocated on the stack, and
      there is no guarantee of the initial value of memory in the stack. So there is no
      need to protect a caller's stack data, but data above the stack frame
      (all of memory) are out of scope and uninitialized, and therefore unsafe to read.

      Let \(\contour : \COMPONENTS \rightarrow \{\HIGHSEC,\LOWSEC\}\) map all addresses
      to \(\HIGHSEC\) (high confidentiality) and all registers to \(\LOWSEC\)
      (low confidentiality). In the style of non-interference, a pair of states \(\mach\)
      and \(\mach'\) are variants with regard to \(\contour\), \(\mach \approx_\contour \mach'\), if:
      \[\forall \component . \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]

      A system enjoys confidentiality in this case if for any initial states \(\mach_0\)
      and \(\mach_0'\) such that \(\mach_0 \approx_\contour \mach_0'\):

      \begin{itemize}
        \item Let \(\mach_0,\mathit{pinit}(\mach_0) \hookrightarrow \MPT\) and \(\mach_0' \hookrightarrow \machT'\)
        \item \(\MPT \sim_P \machT'\)
      \end{itemize}

    \paragraph{Contours}

      Our remaining properties concern the behavior of function calls and require a dynamic treatment
      of confidentiality and integrity. We generalize \(C\) above to a {\it contour} mapping each component
      to a label:

      \[\mathit{label} ::= \{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT,\LOWINT\}\]

      \[\contour \in \CONTOURS ::= \COMPONENTS \rightarrow \mathit{label}\]

      When \(\mpstate{\mach}{\pol}\) is a call state \(\callmap(\mach) = n\) for some \(n\),
      we can construct a contour based on the bounds of the callee's stack frame, as follows:
      \[\mathit{Cof}(\mach,n)(\component) =
      \begin{cases}
        (\HIGHSEC,\HIGHINT) & \text{if } \component \leq \mach(\SP) - n \\
        (\HIGHSEC,\LOWINT) & \text{if } \component \geq \mach(\SP) \\
        (\LOWSEC,\LOWINT) & \text{else} \\
      \end{cases}\]

      Above the stack pointer, memory is high integrity; from the perspective of the callee, that
      memory is not initialized. Below the old stack pointer, memory is high integrity and high
      confidentiality. Registers and the callee's frame are low integrity and low confidentiality.
      These categories inform our integrity and confidentiality properties.

    \paragraph{Eager Integrity}

      A system enjoys eager integrity with regard to \(\callmap\) if for any initial state \(\mach_0\),
      call state \(\mach_c\) where \(\callmap(\mach_c) = n\), and policy state \(\pol_c\), and
      return state \(\mpstate{\mach_r,\pol_r}\), where:
      \[\mach_0,\mathit{pinit}(\mach_0) \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ...
        \mpstate{\mach_r}{\pol_r,\_} | \ret{\mach_c}\]

      it is always the case that:
      \[\forall \component . \contour(\component) = \HIGHINT \rightarrow \mach_c(\component) = \mach_r(\component)\]

    \paragraph{Observable Integrity}

      Observable integrity is concerned with the changes a callee makes to high integrity components,
      and the effects those writes have on the observable behavior of its caller
      {\it after its return}. Given a contour, a call state, and a returned state, we define
      a ``rollback'' function \(\mathit{roll}_I : \CONTOURS \times \MACHS \times \MACHS
      \rightarrow \MACHS\). The rollback creates a state that matches the call
      state on those components that are high integrity in its contour, and matches the return state
      on low integrity components:

      \[\mathit{roll}_I(\contour,\mach_c,\mach_r)(\component) =
      \begin{cases}
        \mach_c(\component) & \contour(\component) = (\HIGHINT,\_) \\
        \mach_r(\component) & \contour(\component) = (\LOWINT,\_) \\
      \end{cases}\]

      Intuitively, the result of a rollback is an idealized state that reflects changes to accessible
      components, but ignores the callee's interference with the caller's data. Then the observable
      behavior of the resulting trace is the standard to which we expect any real trace to conform.

      A system enjoys observable integrity with regard to \(\callmap\) as follows:

      For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpostate{\mach_c}{\pol_c}{\_} ...\)
      and \(\callmap(\mach_c) = n\), if \(\mpstate{\mach_c}{\pol_c} \hookrightarrow ...
        \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\):

          \begin{itemize}
            \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
            \item Let \(\mach_r' = \mathit{roll}_I(\contour,\mach_c,\mach_r)\), \(\mpstate{\mach_r}{\pol_r}
              \hookrightarrow \MPT\), and \(\mach_r' \hookrightarrow \machT'\)
            \item \(\MPT \sim_P \machT'\)
          \end{itemize}

    \paragraph{Eager Confidentiality}

      A system enjoys eager stack confidentiality with regard to \(\callmap\) if, for any initial state
      \(\mach_0\) and call state \(\mach_c\) such that
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ...\) and
      \(\callmap(\mach_c) = n\):

      \begin{itemize}
        \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
        \item For all \(\mach_c'\) such that \(\mach_c \approx_\contour \mach_c'\):
        \item If \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT (\mach_r,p_r,\_) | \ret{\mach_c}\)
          \begin{itemize}
            \item Then there is some \(\mach_r'\) where

              \(\mach_c' \hookrightarrow \machT' (\mach_r',_) | \ret{\mach_c'}\)
            \item \(\mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT')\)
            \item And for all \(\component\), if \(\mach_c(\component) \not = \mach_r(\component)\) or
              \(\mach_c'(\component) \not = \mach_r'(\component)\), \(\mach_r(\component) = \mach_r'(\component)\)
          \end{itemize}
        \item Else, for \(\MPT\) and \(\machT'\) such that \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\)
          and \(\mach_c' \hookrightarrow \machT'\), \(\MPT \sim_P \machT'\)
      \end{itemize}

      \leo{Think more about whether we vary mc or step mc}

    \paragraph{Observable Confidentiality}

      Observable confidentiality is concerned with whether high confidentiality data influences
      observable both behavior during and after a call. We define variants given the generalized notion
      of contours:
      \[\mach \approx_\contour \mach' \triangleq \forall \component .
      \contour(\component) = \LOWSEC \rightarrow \mach(\component) = \mach'(\component)\]
      \leo{Move earlier}

      Data that is high confidentiality within a callee due either to being in its caller's frame
      will become low confidentiality if the callee returns.\leo{??} We face a subtle distinction: 
      a callee treats the contents of its caller's frame as secrets, and they should not influence
      its behavior. Nor should the callee be able to influence behavior after its return based
      on the caller's secrets, for instance by copying them to registers that will later be output.
      But the caller itself may still adjust its behavior based on its own secrets. So our property
      relies on another rollback function to preserve changes made by the caller but remove other
      differences between variant states.

      In this case, a confidentiality rollback takes a contour and four states: two call states
      that will be variants, and two return states. It is not symmetric, regarding \(\mach_c\) and \(\mach_r\)
      as the ``real'' call and return, \(\mach_c'\) as a variant and \(\mach_r'\) as a return that follows it.
      \[\begin{split}
        & \mathit{roll}_C(\contour,\mach_c,\mach_c',\mach_r,\mach_r')(\component) = \\
        & \begin{cases}
          \mach_c(\component) & \text{if } C(\component) = \HIGHSEC \text{ and } \\
                  & \mach_c(\component) = \mach_r(\component) \text{ and }
                    \mach_c'(\component) = \mach_r'(\component) \\
          \mach_r'(\component) & \text{else} \\
        \end{cases}
      \end{split}\]
      \leo{Maybe this should be called unscramble}

      A system enjoys observable confidentiality with respect to \(\callmap\) as follows:

      For any initial state \(\mach_0\), call state \(\mach_c\), and policy state \(\pol_c\) such that
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)} \hookrightarrow ... \mpostate{\mach_c}{\pol_c}{\_} ...\) and
      \(\callmap(\mach_c) = n\):

      \begin{itemize}
        \item Let \(\contour = \mathit{Cof}(\mach_c,n)\)
        \item For all \(\mach_c'\) such that \(\mach_c \approx_C \mach_c'\):
        \item If \(\mpstate{\mach_c}{p_c} \hookrightarrow \MPT_1 \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\)
          \begin{itemize}
            \item Then there is some \(\mach_r'\) where

              \(\mach_c' \hookrightarrow \machT_1' (\mach_r',\_) | \ret{\mach_c'}\)
            \item Let \(\mach_r'' = \mathit{roll}_C(\contour,\mach_c,\mach_r,\mach_c',\mach_r')\),

              \(\mpstate{\mach_r}{p_r} \hookrightarrow \MPT_2\), and \(\mach_r'' \hookrightarrow \machT_2'\)
            \item \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1' \cdot \machT_2'\)
          \end{itemize}
        \item Else, for \(\MPT\) and \(\machT'\) such that \(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\)
          and \(\mach_c' \hookrightarrow \machT'\), \(\MPT \sim_P \machT'\)
      \end{itemize}

      \leo{Need to fix append. Maybe call it join. It should remove the last
        state of left arg}

\end{document}