### 2021-06-15 src/response_formal.txt
Added formal response

** Overview **

** Changes **

1. Give a more thorough discussion of validation. Discuss the feasibility of proofs
   and of randomized testing of more complex properties. Detail our roadmap for
   further validation of different versions of the property.

2. Give an explicit, direct comparison to the work of Skorstengaard et. al. Currently
   we discuss them briefly in our related work section, but our novelty has not been
   obvious, so we can be more thorough in explaining the differences.

3. There are a number of improvements we can make to the presentation. We need to
   gradually introduce the examples and associated intuition (including our visual
   metaphors) as early as possible, starting in the introduction.

4. Draw a clearer distinction between language features that are compatible with our
   model but not assumed, those that are required for the model to function,
   and those that are incompatible.

** Detailed Comments **

* Reviewer A *

We propose above to add further discussion of validation. Here is a rough
outline of how it would cover your concerns.

> ...discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.

These properties would absolutely be amenable to a correctness proof. Such a proof would be a
lot of work in a setting with a full ISA, but there is not reason it should be impossible. The
primary challenge we would predict in such an effort, beyond raw effort, is in determining for a given
enforcement mechanism what invariants it needs to maintain in order to work correctly, and proving that
it maintains them. This is very feasible for a proof of an individual, well-understood enforcement
mechanism, but not for our goal of evaluating several such mechanisms.

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
> 
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

Here we should discuss our roadmap for testing in more detail. At present we are working to implement
testing of observable properties alongside the fixed version of the lazy tagging/clearing micro-policy.
Observational integrity is implemented, but confidentiality is not -- since the lazy policy enforced
normal, stepwise confidentiality this is sufficient to test it. But we did not have this ready before
the deadline. Note that this has been implemented since submission, so we will need to take care not
to describe it as a contribution.

Our other immediate goal is to implement testing of a secondary architecture. We will evaluate a Cheri
implementation of stack safety to demonstrate that our properties are not strictly tied to tag-based
enforcement.

We consider validation of the passing system to be a secondary priority and coroutines tertiary.
The passing system has existing enforcement mechanisms and serves an important role in common
stack use-cases. Coroutines will require the development of an enforcement mechanism (likely
tag-based) to validate.

> There is one topic beyond my evaluation above worth considering: callee-save registers.
> I fear these may be fairly tricky to support in your set-up.  Upon function entry, callee-save
> registers hold data private to the caller that may be passed arbitrarily deeper into the call stack,
> then saved into arbitrary stack locations, but then must be restored before return.
> Have you or the closely related prior work on enforcement mechanisms considered this complexity?
> The invariants are clear enough informally, but the issue seems to add substantial difficulty to
> tracking that a function faithfully implements the calling convention.

Callee-save registers are plausible in our setup, but would require tweaks to the domain system.
They actually provide another use-case for observational properties! Focusing on integrity:
callee-save registers may be overwritten (violating stepwise integrity) but should always be
restored before a return. This behavior corresponds to marking such registers as permanently sealed
in the context of observational integrity -- recall that observational integrity compares sealed
components at the start and end of a callee, so the interim changes are not violations.
(Note as well that the formalization permits domains such as "sealed" to apply to registers as well
as to addresses, we just haven't taken advantage of this ability.)

Confidentiality can be applied similarly. We would permanently treat the stack pointer and
argument registers as "passed", and so exempt from confidentiality. Then, when a callee-save
register is copied to the stack and later restored before return, the register value does
not represent a visible difference in state between the entry and the return.

We had briefly considered this complexity and chosen to avoid it, because it is an extra layer
of detail to apply domains to registers.

> Another minor question: The difference between [nested] call-by-reference and passing
> address-of-local was interesting, but I'm not sure how much it makes sense to argue the
> latter is "beyond the scope of stack safety per se" [line 939] without arguing the same
> of the former.  Perhaps the argument can be made that address-of-local can be put in the
> heap while a by-reference parameter cannot and since you don't guarantee heap safety,
> you shouldn't try to do anything useful with address-of-local?

Yes, this is the sort of issue that makes address-of-local hard to handle without heap safety.
More broadly the issue is one of assuming intent on the part of the code. We tend to
assume that an address-of-local should be treated similarly to the address of a heap object,
but with a different lifetime. For instance, if a heap model treats heap pointers like capabilities,
we would treat stack pointers the same. In heap model that permits pointer forging (restricted to
within compartments, say,) we would expect to be able to forge pointers to address-taken locals
in the same circumstances. By contrast, the expected behavior of stack allocated arguments is
closely tied to that of the stack itself.

In the longer term we would like to extend our model with a heap, and then the interplay between
heap-derived and stack-derived pointers (with similar behaviors but different lifespans) will
be much more interesting.

* Reviewer B *

> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.^M

We improve upon Skorstengaard et. al. in the following ways.
1. We capture the dynamic nesting of security concerns implicit in stack-based reasoning
2. We therefore do not assume a static division of trusted and untrusted code

Taken together, these distinctions give the practical advantage that we can reason about stack safety
in contexts where a buggy function is compromised and used to attack other instances of the same function.

3. We clearly state the property in terms of a machine's behavior, without appealing to the reader's
understanding of a specific abstract machine
4. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
machine being architecturally similar to the concrete one (beyond the specific call and return semantics
that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
abstract machine. Our properties apply directly to a wide range of architectures.

We also differ from Skorstengaard et. al. in our focus. We use the language of formal security properties
to apply the established intuitions and formal techniques of non-interference. This demonstrates that stack
safety is expressible as a formal security property.

> Q2. In L117-121 you state certain assumptions about the language semantics. It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply. For example, does a language with a heap allow you to break stack safety? What about a language with delimited continuations that manipulate the stack?^M

The section should be broken into assumptions that are necessarily for our properties, and assumptions
that we explicitly do not make. Our stack safety model applies regardless of how a language uses non-stack,
non-code memory, including to languages with a heap. They apply to any layout of memory as long as it is
known in advance.

Delimited continuations are not supported by our base model, so a language that has them would violate
stack safety. We extend the base model with argument passing on the stack to demonstrate how it can
accommodate a programming language feature that shares stack data between callers and callees. A
similar extension to delimited continuations would be more complex.

> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et al (L289-291)? What are the implications of that?

We want our properties to be applicable even to low-level code that may not keep function
footprints isolated from one another. Consider a program that mixes C code with hand-written
assembly -- the assembly may make calls and returns, which should be protected by stack safety,
as well as arbitrary jumps, which should not.

In Skorstengaard, an untrusted adversary can exhibit such behavior, but only trusted code can
expect its data to be protected. We extend this protection to all callers, even those that
exhibit unusual control flow.

* Reviewer C *

> I might have missed something important. Are there any applications of the proposed formalization?

A formal definition of a type of security is valuable in its own right. The main application
is identifying bugs in real world enforcement mechanisms -- without such a definition, we
can't necessarily agree that a given behavior is a bug. It gives us the ability to reason
clearly about what is gained from applying a given enforcement mechanism to a system. Since
many enforcement mechanisms are quite costly in speed, dollars, and silicon, it is important
to understand precisely what they buy us.