### 2021-06-16 submission2/response_formal.txt
Polishing on first half of response

# Overview

We thank the reviewers for their comments.

The reviews identify some areas that the paper fails to explain
adequately -- in particular, "framing" questions: How does the work improve
on existing formalizations of stack safety? Do we have a plan to validate
the model further?  Can our work be extended to cover more complex uses of
the stack?

In the next section, we outline the changes we plan to make in the paper to
clarify these questions.  We discuss the issues in more detail in the third
section.

## Changes

1. We will give a more thorough discussion of validation, discuss the
   feasibility of proofs and of randomized testing of more complex
   properties, and detail our roadmap for further validation of different
   versions of the property.

2. We will give a more explicit and direct comparison to the work of
   Skorstengaard et. al. Currently we discuss it briefly in our related work
   section, but the novelty of our work does not come through clearly
   enough; we will explain the differences more thoroughly.

3. We will improve the presentation as suggested. In particular, we will
   gradually introduce the examples and associated intuitions (including our
   visual metaphors) as early as possible, starting in the introduction.

4. We will more clearly call out common language features, such as the heap,
   that are compatible with our treatment of stack safety.

5. [SNA: Do we want to promise to add callee-saves registers? I wasn't going to
   at first, but actually there's a good story here. Something like:] [BCP: If
   we think we can really do this, we should say so!]
   We will add a discussion of how to extend the model to a callee-saves
   registers convention, which demonstrates a concrete use for the observational
   properties.
   [BCP: Doesn't the current draft also demonstrate a use for the observational
   properties?]

# Detailed Comments

## Reviewer A

We will extend the paper with a more detailed discussion of our roadmap for
verification and testing and the issues that we expect to arise. Here is a
rough outline of how it would cover the concerns raised in the review.

> ...discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.

The formal security properties we define would absolutely be amenable to a
correctness proof. Such a proof would be considerable work to carry out for
a full ISA like RISC-V, but we see no fundamental barriers.  We chose to
concentrate on testing because it is better aligned with our immediate goal
of evaluating real enforcement mechanisms for real machine architectures.

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
>
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

We are working to implement testing of observable properties and to apply
this to the putatively fixed version of the lazy tagging/clearing
micro-policy.  At the moment, testing of observational integrity is
implemented; observational confidentiality testing is not, but -- since the
lazy policy enforces normal, stepwise confidentiality -- this is actually
sufficient to test it. [APT: Do we have the results of these tests?][BCP: +1
-- if we could actually say something about results it would be fantastic!]
But we did not have this ready before the deadline. [BCP: IIRC, we have not
been forbidden from mentioning new results that we would include in the
final paper if accepted, and IMO there is little risk of annoying the
reviewers if we mention them in response to direct criticisms or questions.]

Our other short-term plan is to implement testing of a second enforcement
mechanism -- a Cheri implementation of stack safety -- to demonstrate that
our properties are not strictly tied to tag-based enforcement. We are
currently working on the QuickChick generators for Cheri programs, which
have a different calling convention (a more complicated one, which is easier
to get wrong!).  [BCP: We should probably add something like: "But this
effort falls outside the scope of the current paper" to make clear that we
are not promising it.]

Validation of the argument-passing and coroutine models are longer-term
projects.  The passing system has existing enforcement mechanisms and serves
an important role in common stack use-cases. Coroutines will require the
development of an enforcement mechanism (e.g., one based on tags) to
validate.

> There is one topic beyond my evaluation above worth considering:
> callee-save registers.  I fear these may be fairly tricky to support in
> your set-up.  Upon function entry, callee-save registers hold data private
> to the caller that may be passed arbitrarily deeper into the call stack,
> then saved into arbitrary stack locations, but then must be restored
> before return.  Have you or the closely related prior work on enforcement
> mechanisms considered this complexity?  The invariants are clear enough
> informally, but the issue seems to add substantial difficulty to tracking
> that a function faithfully implements the calling convention.

Callee-save registers are compatible with our approach, using a version of
the sharing model with its context initialization and updates modified to
apply appropriate domains to registers.  Indeed, they actually provide
another interesting use-case for the "observational" variants of our
properties. Focusing on integrity: callee-save registers may be overwritten
(violating stepwise integrity) but should always be restored before a
return. This requirement corresponds to observational integrity with the
callee-save registers sealed. (Recall that observational integrity compares
sealed components at the start and end of a callee's execution, so the
interim changes are not violations.)

Confidentiality for callee-saves is similar. A callee-save register should
be secret, but in observational confidentiality, if it is copied to the
stack and then restored before return, the property will allow the original
register to be read normally after return.  But the property will not allow
it to be leaked in any other way.

We had considered adding this complexity and chosen to avoid it, because it
is an extra layer of detail to apply domains to registers.  [BCP: Don't
understand this.  Maybe just drop it and say we will add this discussion to
the paper?  Or do we even want to say that we think we can extend the model
and our testing framework to callee-saves before the camera-ready?... My
sense is that this is too much work, but either way we should say (again)
what we plan to do.]

It is worth noting that the main prior work in this area by Skorstengaard et
al. relies [BCP: fundamentally?] on a calling convention which clears all
non-argument registers before calls and all non-return registers before
returns.

> Another minor question: The difference between [nested] call-by-reference
> and passing address-of-local was interesting, but I'm not sure how much it
> makes sense to argue the latter is "beyond the scope of stack safety per
> se" [line 939] without arguing the same of the former.  Perhaps the
> argument can be made that address-of-local can be put in the heap while a
> by-reference parameter cannot and since you don't guarantee heap safety,
> you shouldn't try to do anything useful with address-of-local?

Yes, this was exactly our thinking.
We know what constitutes legal access to stack-allocated arguments, but for explicit pointers
this is much less clear, and depends on the intent of the programmer. We believe
an address-of-local should be treated similarly to the address of a heap object,
but with a different lifetime. For instance, if a heap safety model treats heap pointers like capabilities
(possession implies access rights), we would treat stack pointers the same.
Similarly, if the heap model permits pointer forging (restricted to
within compartments, say) we would expect to be able to forge pointers to address-taken locals
in the same circumstances.

In the longer term we would like to extend our model with a heap, in the style of
Azevedo de Amorim et al.'s "The Meaning of Memory Safety," which is phrased in similar
terms to our model and should be compatible. Then the interplay between heap-derived and
stack-derived pointers (with similar behaviors but different lifespans) will
be much more interesting.

* Reviewer B *

> Q1. How does this work improve upon Skorstengaard et al? Neither the introduction nor related work section give a satisfying answer to this.^M

We improve upon Skorstengaard et. al. in the following ways.
1. We capture a desireable security characteristic of dynamically nested calls, that every caller is safe from its callee
2. We therefore do not assume a static division of trusted and untrusted code

Specifically, Skorstengaard et. al. divide the world into trusted functions, which feature
additional dynamic constraints on their behavior, and for which their properties hold,
and adversaries, which may violate the dynamic "reasonableness" constraints. Suppose we
have a single function that recursively traverses a linked list, loading data into local
memory. What if it has a bug, and certain data cause it to violate the reasonableness condition?
In Skorstengaard's model, it must be an adversary, and activations lower in the stack are not
guaranteed protection. In our model, only the badly behaved activation is an attacker, and its
caller(s) are guaranteed protection.

3. Skorstengaard's approach to proving safety via fully-abstract overlay semantics relies on the abstract
machine being architecturally similar to the concrete one (beyond the specific call and return semantics
that distinguish them). Therefore it is less portable: each architecture it applies to wants its own
abstract machine. Our properties should apply directly to a wide range of
architectures.

We also differ from Skorstengaard et. al. in our focus. We use the language of formal security properties
to apply the established intuitions and formal techniques of non-interference. This demonstrates that stack
safety is expressible as a formal security property.

> Q2. In L117-121 you state certain assumptions about the language semantics. It is unclear to me if these assumptions restrict the languages to which your stack safety definition apply. For example, does a language with a heap allow you to break stack safety? What about a language with delimited continuations that manipulate the stack?^M

Our stack safety model applies regardless of how a language uses non-stack,
non-code memory, so it works for languages with a heap. It applies to any layout of memory as long as that is
known statically.

Delimited continuations are not supported by our base model, so a language that has them would violate
stack safety. We extend the base model with argument passing on the stack to demonstrate how it can
accommodate a programming language feature that shares stack data between callers and callees. A
similar extension to delimited continuations would be more complex.

This distinction, between a feature that could be added without changing the properties and one that
would require enhanced properties, is important, and we will clarify in section 2 which features fall
into which category.

> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et al (L289-291)? What are the implications of that?

We want our properties to be applicable even to low-level code that may not keep function
footprints isolated from one another. Consider a program that mixes C code with hand-written
assembly -- the assembly may make calls and returns, which should be protected by stack safety,
as well as arbitrary jumps, which should not.

* Reviewer C *

> - It is a bit difficult to assess the merits of this work. Yes, the
>   stack safety property is important for security, but what new results
>   or new capabilities does the proposed definition bring in?

A formal definition of a type of security is valuable in its own right. It tells us whether
any given behavior is a violation, and we can therefore use it to identify bugs. In our case
study we identify a previously unknown bug in Roessler and Dehon's proposed "lazy tagging" micro-policy,
and also use randomized testing to increase confidence that their "depth isolation" micro-policy
correctly implements stack safety.

More broadly, it gives us the ability to reason
clearly about what is gained from applying a given enforcement mechanism to a system. Since
many enforcement mechanisms are quite costly in speed, dollars, and silicon, it is important
to understand precisely what they buy us.

>   The paper does not talk about data flow
>   across heap and stack, which is an important element for many
>   real-world memory corruption attacks.

We plan to combine our stack model with a heap model in the style of Acevedo de Amorim et al.
as future work.

> - The paper is titled "stack safety is a security property". I don't
>   see why this is surprising, or what's the novelty here. Memory safety
>   is almost always connected with security properties.

We agree that the title is confusing and should be changed. It refers to a concept in
theoretical security: a formal notion of security property as a mathematical description of
the possible behaviors of a system. In that context, it was not actually obvious that stack
safety could be defined in these terms, but now we have done so. That said, the title should
not require readers to understand this specific terminology to be comprehensible.