### 2021-06-17 submission2/response_formal.txt
one last bit of rewording and reordering a few bullets

# Overview

We thank the reviewers for their comments.

The reviews identify some areas that the paper fails to explain
adequately -- in particular, "framing" questions: How does the work improve
on existing formalizations of stack safety? Do we have a plan to validate
the model further?  Can our work be extended to cover more complex uses of
the stack?

In the next section, we outline the changes we plan to make in the paper to
clarify these questions.  We discuss the issues in more detail in the third
section.

## Changes

1. We will give a more thorough discussion of validation, discuss the
   feasibility of proofs and of randomized testing of more complex
   properties, and detail our roadmap for further validation of different
   versions of the property. For details, see the comments responding
   to reviewer A.

2. We will give a more explicit and direct comparison to the work of
   Skorstengaard et. al. Currently we discuss it briefly in our related work
   section, but the novelty of our work does not come through clearly
   enough; we will explain the differences more thoroughly, all in one place.
   For details, see the comments responding to reviewer B.

3. We will improve the presentation as suggested. In particular, we will
   gradually introduce the examples and associated intuitions (including our
   visual metaphors) as early as possible, starting in the introduction.

4. We will more clearly call out common language features, such as the heap,
   that are compatible with our treatment of stack safety. For details,
   see the comments responding to reviewer B.

5. We will add a discussion of how to extend the model to a callee-saves
   registers convention, which demonstrates a second concrete use for 
   the observational properties. Previously we were concerned about space
   and explanatory complexity, but in retrospect we see that discussing
   register protection would paint a more complete picture.
   For details, see the comments responding to reviewer A.

# Detailed Comments

## Reviewer A

We will extend the paper with a more detailed discussion of our roadmap for
verification and testing and the issues that we expect to arise. Here is a
rough outline of how it would cover the concerns raised in the review.

> ...discussion on whether it is feasible/desirable/in-progress/etc. to prove enforcement mechanisms correct.

Correctness of an enforcement mechanism with respect to our properties
is certainly amenable to formal proof. Such a proof would be
considerable work to carry out for a full ISA like RISC-V, but we see
no fundamental barriers.  We chose to concentrate on testing because
it is better aligned with our immediate goal of evaluating real
enforcement mechanisms for real machine architectures.

> * The testing for the lazy tagging/clearing found mistakes but the "potential fix [is] itself a prime candidate for future testing" [line 868]
>
> * Any validation for the lazier observability definition (this is listed as future work.)^M
>
> * Any validation for the coroutine work.

[ We have since tested lazy tagging/clearing and found it to pass the
  newly implemented observational integrity property as well as normal,
  stepwise confidentiality. But we did not have this ready before the
  deadline.
| We are working to implement testing of observational properties and to apply
  this to the putatively fixed version of the lazy tagging/clearing
  micro-policy. ]

Our other short-term plan is to implement testing of a second enforcement
mechanism -- a Cheri implementation of stack safety -- to demonstrate that
our properties are not strictly tied to tag-based enforcement. We are
currently working on QuickChick generators for Cheri programs, which
have a different calling convention (a more complicated one, which is easier
to get wrong!). This effort is beyond the scope of this paper, but we will
sketch it as future work.

Validation of the argument-passing and coroutine models are longer-term
projects.  The argument-passing model has existing enforcement mechanisms and serves
an important role in common stack use cases. Coroutines will require the
development of an enforcement mechanism (e.g., one based on tags) to
validate.

> There is one topic beyond my evaluation above worth considering:
> callee-save registers.  I fear these may be fairly tricky to support in
> your set-up.  Upon function entry, callee-save registers hold data private
> to the caller that may be passed arbitrarily deeper into the call stack,
> then saved into arbitrary stack locations, but then must be restored
> before return.  Have you or the closely related prior work on enforcement
> mechanisms considered this complexity?  The invariants are clear enough
> informally, but the issue seems to add substantial difficulty to tracking
> that a function faithfully implements the calling convention.

Callee-save registers are supported by our approach, using a
version of the sharing model with its context initialization and
updates modified to apply appropriate domains to registers as well as
stack locations.  Indeed, they actually provide another interesting
use case for the "observational" variants of our properties. Focusing
on integrity: callee-save registers may be overwritten (violating
stepwise integrity) but should always be restored before a
return. This requirement corresponds to observational integrity with
the callee-save registers sealed. (Recall that observational integrity
compares sealed components at the start and end of a callee's
execution, so the interim changes are not violations.)

Confidentiality for callee-saves is similar. A callee-save register should
be secret, but in observational confidentiality, if it is copied to the
stack and then restored before return, the property will allow the original
register to be read normally after return.  But the property will not allow
it to be leaked in any other way.

It is worth noting that the main prior work in this area by Skorstengaard et
al. relies fundamentally on a calling convention that clears all
non-argument registers before calls and all non-return registers before
returns.

We will add this discussion to the paper, where previously we had omitted
it due to concerns for space and for the complexity of the explanations.
But it both demonstrates the model's flexibility and ties in neatly to
observational properties, so it is worth including.

> Another minor question: The difference between [nested] call-by-reference
> and passing address-of-local was interesting, but I'm not sure how much it
> makes sense to argue the latter is "beyond the scope of stack safety per
> se" [line 939] without arguing the same of the former.  Perhaps the
> argument can be made that address-of-local can be put in the heap while a
> by-reference parameter cannot and since you don't guarantee heap safety,
> you shouldn't try to do anything useful with address-of-local?

Yes, this was exactly our thinking.  We know what constitutes legal access
to stack-allocated arguments, but for explicit pointers this is much less
clear, and depends on the intent of the programmer. We believe an
address-of-local should be treated similarly to the address of a heap
object, but with a different lifetime. For instance, if a heap safety model
treats heap pointers like capabilities (possession implies access rights),
we would treat stack pointers the same.  Similarly, if the heap model
permits pointer forging (restricted to within compartments, say) we would
expect to be able to forge pointers to address-taken locals in the same
circumstances.

In the longer term we would like to extend our model with a heap, in
the style of Azevedo de Amorim et al. [1] which is phrased in similar
terms to our model and should be compatible. In this setting the
interplay between heap-derived and stack-derived pointers (with
similar behaviors but different lifespans) will be much more
interesting.

## Reviewer B

> Q1. How does this work improve upon Skorstengaard et al? Neither the
> introduction nor related work section give a satisfying answer to this.

We improve upon Skorstengaard et. al. in the following ways.

1. We capture a desireable security characteristic of dynamically nested
   calls: that EVERY caller should be protected from its callees.

   By contrast, Skorstengaard et. al. statically divide the world into
   trusted functions, which feature additional dynamic "reasonableness"
   constraints on their behavior and for which their properties hold,
   and adversary functions, which may violate those constraints but
   are not guaranteed safety.

   The dynamic nature of this boundary makes it an issue with 
   functions that cannot be statically proven safe, but whose data
   still need to be protected when they are behaving safely.
   For a contrived example, consider a function that stores some
   private state, but takes an argument "unsafe" and, if called with
   "unsafe" set to 1, violates the reasonableness constraints. In
   the StkTokens model, we can never guarantee the safety of the
   data even if called with "unsafe" set to 0.

   We will expand on such an example in the revised version.

2. Skorstengaard's approach to proving safety via fully abstract overlay
   semantics relies on the abstract machine being architecturally similar to
   the concrete one (beyond the specific call and return semantics that
   distinguish them). This makes it less portable: each architecture it
   applies to needs its own abstract machine. Our properties should apply
   directly to a wide range of architectures.

3. We also differ from Skorstengaard et. al. in the nature of our
   safety definition.  We use the language of formal security
   properties to apply well-established concepts and formal techniques
   of non-interference. Thus, we define our properties in terms of
   traces of memory and register states, rather than in terms of an
   ad-hoc abstract machine that is asserted to be "stack safe by construction."
   Which approach is better may be a matter of taste, but we believe that our
   alternative has the merit of explicitly relating stack safety to existing
   notions of security.

> Q2. In L117-121 you state certain assumptions about the language
> semantics. It is unclear to me if these assumptions restrict the languages
> to which your stack safety definition apply. For example, does a language
> with a heap allow you to break stack safety? What about a language with
> delimited continuations that manipulate the stack?

Our stack safety model applies regardless of how a language uses non-stack,
non-code memory, so it works for languages with a heap.  It also applies to
any (statically known) memory layout.

Delimited continuations are not supported by our base model, so a language
that has them would violate stack safety. We do extend the base model with
argument passing on the stack to demonstrate how it can accommodate a
language feature that shares stack data between callers and callees, but a
similar extension to delimited continuations would be more complex.

This distinction -- between features that could be added without changing
the properties and ones that would require enhanced properties -- is
important, and we will clarify in section 2 which features fall into which
category.

Some features that can be added/changed without changing the properties:
- a heap, as stated
- calling convention choices, like whether or not to have a frame pointer
- the machine's ISA

Features that cannot be supported without changes to the model:
- address-taken locals (would be similar to argument passing, but
  would benefit from more sophisticated, heap-like memory models)
- exceptions (needs annotations of throw/catch points)
- continuations (would be similar to argument passing)
- runtime code generation (needs dynamic code annotations)
- parallelism (needs fundamental changes to step, trace models)

> Q3. Why don't you assume well-bracketed control flow like Skorstengaard et
> al (L289-291)? What are the implications of that?

Skorstengaard et al. don't directly assume well-bracketed control
flow, they enforce it. But the terms in which they define it differ
from ours. Their version assumes a code layout where each function has
an isolated footprint and they don't permit unrestricted jumps between
footprints.  Our definitions are sensible without that assumption.

## Reviewer C

> - It is a bit difficult to assess the merits of this work. Yes, the
>   stack safety property is important for security, but what new results
>   or new capabilities does the proposed definition bring in?

First, we contend that a formal definition of stack safety is valuable in
its own right, as a way of shifting the discussion away from specific
examples of bad behavior that must be prevented and toward general
mechanisms that ensure good behavior.  A formal definition gives us the
ability to reason clearly about what is gained from applying a given
enforcement mechanism to a system. Since many enforcement mechanisms are
quite costly in speed, dollars, and/or silicon, it is important to
understand precisely what they buy us.

More immediately, we can use a formal definition like this to test proposed
enforecement mechanisms for violations of the policy.  In our case study, we
identify a previously unknown bug in Roessler and Dehon's proposed "lazy
tagging" micro-policy, and we use randomized testing to increase confidence
that their "depth isolation" micro-policy correctly implements stack safety.

>   The paper does not talk about data flow
>   across heap and stack, which is an important element for many
>   real-world memory corruption attacks.

Agreed!  We plan to combine our stack model with a heap model in the style
of Azevedo de Amorim et al. [1] as future work.  Both models are phrased in the
conceptual terminology of language-based security, so we have reason to
believe that they will fit together cleanly.

> - The paper is titled "stack safety is a security property". I don't
>   see why this is surprising, or what's the novelty here. Memory safety
>   is almost always connected with security properties.

In hindsight, our choice of title was confusing: We did not just mean that
stack safety impacts security (which is obvious), but that it can usefully
be formulated using (a refinement of) the same formal terminology and
mathematical definitions -- confidentiality, integrity, noninterference,
etc. -- that have been used for decades to reason about security.

We plan to change our title to:
"Formalizing Stack Safety as a Security Property"

----------------

[1] A. Azevedo de Amorim, C. Hritcu, and B. Pierce, "The Meaning of Memory Safety,"
Proc. Principles of Security and Trust, 2018.